---
title: "Clustering"
author: 'Noémie Wellinger'
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(blockcluster)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(leaflet)
library(viridisLite)  # For viridis color palette
library(RColorBrewer)
```

```{r}
# load data
all_data_hourly <- readRDS("../data/final_predictors/all_data_hourly.rds") 
all_data_daily_night <- readRDS("../data/final_predictors/all_data_daily.rds") |>
  filter(daynight == "n") # only night data

# Setup
cluster_variable <- "UHI"
```

```{r}
# Select data
data_to_cluster <- all_data_hourly |>
  select(time, Log_NR, UHI) |>
   pivot_wider(names_from = Log_NR, values_from = UHI)

# Set time_hour as row names
data_to_cluster <- as.data.frame(data_to_cluster)
rownames(data_to_cluster) <- data_to_cluster$time
data_to_cluster$time <- NULL
#remove all rows that contain NA
data_to_cluster <- data_to_cluster[complete.cases(data_to_cluster),]

# Convert to matrix
matrix_to_cluster <- as.matrix(data_to_cluster)

# Ensure row names are preserved
rownames(matrix_to_cluster) <- rownames(data_to_cluster)
```


```{r}
# Define the number of clusters (tune this parameter)
num_row_clusters <- 4  # Time periods 
num_col_clusters <- 7  # Stations

set.seed(124)  # For reproducibility
# Run co-clustering
result <- coclusterContinuous(matrix_to_cluster, nbcocluster = c(num_row_clusters, num_col_clusters))
plot(result)
# save the plot
ggsave(paste0("../results/Clusters/CoCluster_", timestep, "_", num_col_clusters, "S_", num_row_clusters, "T.png"), cocluster_plot)
summary(result)
```




```{r}
print(result@rowclass)  # Time clusters
print(result@colclass)  # Station clusters
```

```{r}
# Attach the cluster classes to the original data
# Convert to data frames for merging

# Extract row (time) and column (station) cluster assignments
time_cluster_df <- data.frame(time = rownames(matrix_to_cluster), time_cluster = result@rowclass)
time_cluster_df$time <- as.POSIXct(time_cluster_df$time, format = "%Y-%m-%d %H:%M:%S", tz = "Europe/Zurich")
station_cluster_df <- data.frame(Log_NR = colnames(matrix_to_cluster), station_cluster = result@colclass)

# Merge time clusters with original data
all_data_hourly_clusters <- all_data_hourly |>
  left_join(time_cluster_df, by = "time") |>
  left_join(station_cluster_df, by = "Log_NR")

# Prepare for Mapping
# group by time_cluster and Log_NR and summarize
time_clusters_UHI <- all_data_hourly_clusters %>%
  group_by(time_cluster, Log_NR) %>%
  summarize(UHI = mean(UHI, na.rm = TRUE)) |>
  left_join(meta, by = "Log_NR")
```


```{r}
#hourly
cluster_colors <- c('#fc8d62','#8da0cb','#a6d854','#e78ac3','#66c2a5','#ffd92f')

# Plot the hour of day (y axis) and date (y axis) for each data point in data_long_hourly_clusters
all_data_hourly_clusters |>
  mutate(date = as.Date(time),
         hour = hour(time)) |>
  ggplot(aes(x = date, y = hour, fill = factor(time_cluster))) +
  geom_tile() +
  theme_minimal() +
  labs(title = paste(cluster_variable, "Time Cluster Assignments"), fill = "Time Cluster") +
  scale_fill_manual(values = cluster_colors)

# same plot just for months August and September
all_data_hourly_clusters |>
  mutate(date = as.Date(time),
         hour = hour(time)) |>
  filter(date >= "2024-08-01" & date <= "2024-09-30") |>
  ggplot() +
  geom_tile(aes(x = date, y = hour, fill = factor(time_cluster))) +
  theme_minimal() +
  labs(title = paste(cluster_variable, "Time Cluster Assignments (August and September)"), fill = "Time Cluster") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = cluster_colors)
  # plot a line with the average temperature for each day from daily_summary
  #geom_line(data = daily_summary, aes(x = time_day, y = T_mean_day, color = "red"), size = 1) 

ggsave(paste0("../results/Clusters/TimeCluster_timeseries", timestep, "_", cluster_variable, ".png"), width = 6, height = 4, dpi = 300)
```

# distribution of variables, hourly
```{r}
# map clusters for different meteorological variables
all_data_hourly_clusters |>
  select(time, Log_NR, UHI, time_cluster, station_cluster,
         air_temp, humidity, wind_vel, wind_dir_class, rad_1d, pressure, precip_1d)


# Define the variables to normalize
vars_to_normalize <- c("air_temp", "humidity", "wind_vel", "rad_1d", "pressure", "precip_1d")

# Normalize variables to 0–1 range
normalized_data <- all_data_hourly_clusters |>
  mutate(across(all_of(vars_to_normalize), ~ (. - min(., na.rm = TRUE)) / 
                                            (max(., na.rm = TRUE) - min(., na.rm = TRUE)))) |>
  select(time_cluster, all_of(vars_to_normalize)) |>
  pivot_longer(cols = all_of(vars_to_normalize), 
               names_to = "variable", values_to = "value") |>
  filter(!is.na(value))

# Plot
normalized_data |>
  mutate(time_cluster = factor(time_cluster, levels = unique(time_cluster))) |>
  group_by(time_cluster) |>
ggplot(aes(x = time_cluster, y = value, fill = time_cluster)) +
  geom_boxplot(outlier.size = 0.5) +
  facet_wrap(~variable, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Time Cluster", y = "Normalized Value", title = "Distribution of Normalized Variables by Time Cluster") +
  theme_minimal() +
  theme(legend.position = "none") +
  # arrange the time cluster in the order 0, 3, 1, 2, NA on the x axis
  scale_x_discrete(limits = c("2", "3", "0", "1", NA)) +
  scale_fill_manual(values = c("0" = "#fc8d62", "3" = "#e78ac3", "1" = "#8da0cb", "2" = "#a6d854"))

ggsave(paste0("../results/Clusters/Normalized_Variables_by_Time_Cluster_", timestep, ".png"), width = 6, height = 4, dpi = 400)
```




2. Night mean clusters
```{r}
# load data
all_data_daily_night <- readRDS("../data/final_predictors/all_data_daily.rds") |>
  filter(daynight == "n") # only night data

# Setup
cluster_variable <- "UHI"
timestep <- "night"

# Select data
data_to_cluster <- all_data_daily_night |>
  select(time, Log_NR, UHI) |>
   pivot_wider(names_from = Log_NR, values_from = UHI)

# Set time_hour as row names
data_to_cluster <- as.data.frame(data_to_cluster)
rownames(data_to_cluster) <- data_to_cluster$time
data_to_cluster$time <- NULL
#remove all rows that contain NA
data_to_cluster <- data_to_cluster[complete.cases(data_to_cluster),]

# Convert to matrix
matrix_to_cluster <- as.matrix(data_to_cluster)

# Ensure row names are preserved
rownames(matrix_to_cluster) <- rownames(data_to_cluster)
```


```{r}
# Define the number of clusters (tune this parameter)
num_row_clusters <- 4  # Time periods 
num_col_clusters <- 7  # Stations

set.seed(124)  # For reproducibility
# Run co-clustering
result <- coclusterContinuous(matrix_to_cluster, nbcocluster = c(num_row_clusters, num_col_clusters))
plot(result)

summary(result)
plot(result, type = "distribution")
# save as png
#save.png(p, paste0("../results/Clusters/CoCluster_distribution_", timestep, "_", num_col_clusters, "S_", num_row_clusters, "T.png"), width = 800, height = 600)
```




```{r}
print(result@rowclass)  # Time clusters
print(result@colclass)  # Station clusters
```

```{r}
# Attach the cluster classes to the original data
# Convert to data frames for merging

# Extract row (time) and column (station) cluster assignments
time_cluster_df <- data.frame(time = rownames(matrix_to_cluster), time_cluster = as.factor(result@rowclass))
time_cluster_df$time <- as.POSIXct(time_cluster_df$time, format = "%Y-%m-%d", tz = "Europe/Zurich")
station_cluster_df <- data.frame(Log_NR = colnames(matrix_to_cluster), station_cluster = as.factor(result@colclass))

# Merge time clusters with original data
all_data_daily_night_clusters <- all_data_daily_night |>
  left_join(time_cluster_df, by = "time") |>
  left_join(station_cluster_df, by = "Log_NR")

# Prepare for mapping
# group by time_cluster and Log_NR and summarize
time_clusters_UHI <- all_data_daily_night_clusters |>
  group_by(time_cluster, Log_NR) |>
  summarize(UHI = mean(UHI, na.rm = TRUE)) |>
  left_join(meta, by = "Log_NR")

station_clusters_UHI <- all_data_daily_night_clusters |>
  group_by(Log_NR) |>
  summarize(OST_CHTOPO = mean(OST_CHTOPO, na.rm = TRUE),
            NORD_CHTOPO = mean(NORD_CHTOPO, na.rm = TRUE),
            station_cluster = first(station_cluster),
            UHI = mean(UHI, na.rm = TRUE)) 
```

# daily
```{r}
#daily
cluster_colors <- c('#a6d854','#fc8d62','#8da0cb','#e78ac3','#66c2a5','#ffd92f')

# Plot the hour of day (y axis) and date (y axis) for each data point in data_long_hourly_clusters
all_data_daily_night_clusters |>
  mutate(date = as.Date(time),
         hour = 0) |>
  ggplot(aes(x = date, y = hour, fill = factor(time_cluster))) +
  geom_tile() +
  theme_minimal() +
  labs(title = paste(cluster_variable, "Time Cluster Assignments"), fill = "Time Cluster") +
  scale_fill_manual(values = cluster_colors)

# same plot just for months August and September
all_data_daily_night_clusters |>
  mutate(date = as.Date(time),
         hour = 0) |>
  filter(date >= "2024-08-01" & date <= "2024-09-30") |>
  ggplot() +
  geom_tile(aes(x = date, y = hour, fill = factor(time_cluster))) +
  theme_minimal() +
  labs(title = paste(cluster_variable, "Time Cluster Assignments (August and September)"), fill = "Time Cluster") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = cluster_colors)
  # plot a line with the average temperature for each day from daily_summary
  #geom_line(data = daily_summary, aes(x = time_day, y = T_mean_day, color = "red"), size = 1) 
ggsave(paste0("../results/Clusters/TimeCluster_timeseries", timestep, "_", cluster_variable, ".png"), width = 6, height = 4, dpi = 300)

```

# Distribution of variables, daily
```{r}
# map clusters for different meteorological variables
all_data_daily_night_clusters |>
  select(time, Log_NR, UHI, time_cluster, station_cluster,
         air_temp, humidity, wind_vel, wind_dir_class, rad_1d, pressure, precip_1d)


# Define the variables to normalize
vars_to_normalize <- c("air_temp", "humidity", "wind_vel", "rad_1d", "pressure", "precip_1d")

# Normalize variables to 0–1 range
normalized_data <- all_data_daily_night_clusters |>
  mutate(across(all_of(vars_to_normalize), ~ (. - min(., na.rm = TRUE)) / 
                                            (max(., na.rm = TRUE) - min(., na.rm = TRUE)))) |>
  select(time_cluster, all_of(vars_to_normalize)) |>
  pivot_longer(cols = all_of(vars_to_normalize), 
               names_to = "variable", values_to = "value") |>
  filter(!is.na(value))

# Plot
normalized_data |>
  mutate(time_cluster = factor(time_cluster, levels = unique(time_cluster))) |>
  group_by(time_cluster) |>
ggplot(aes(x = time_cluster, y = value, fill = time_cluster)) +
  geom_boxplot(outlier.size = 0.5) +
  facet_wrap(~variable, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Time Cluster", y = "Normalized Value", title = "Distribution of Normalized Variables by Time Cluster") +
  theme_minimal() +
  theme(legend.position = "none") +
  # arrange the time cluster in the order 0, 3, 1, 2, NA on the x axis
  scale_x_discrete(limits = c("0", "3", "1", "2", NA)) +
  scale_fill_manual(values = c("0" = "#a6d854", "3" = "#e78ac3", "1" = "#fc8d62", "2" = "#8da0cb"))

ggsave(paste0("../results/Clusters/Normalized_Variables_by_Time_Cluster_", timestep, ".png"), width = 6, height = 4, dpi = 400)

```





# Mapping (for daily and hourly)
```{r}
# for all the values of time_cluster, plot the UHI_mean
for(i in 0:3){
  p <- time_clusters_UHI |>
    filter(time_cluster == i) |>
    ggplot(aes(x = OST_CHTOPO, y = NORD_CHTOPO, fill = UHI)) +
      geom_point(      
        color = "black",
        shape = 21,
        size = 4,
        stroke = 0.4) +
      labs(title = paste(timestep, "Time Cluster N°", i), 
           fill = "UHI",
           x = "Longitude", y = "Latitude") +
      theme(legend.position = "bottom") +
      # Use color palette RdYlBu with midpoint at 0
      #scale_color_gradient2(low = "blue", mid = "beige", high = "red", midpoint = 0, limits= c(-5, 5)) +
      scale_fill_viridis_c(name = "UHI [K]", limits= c(-1, 5.1)) +
      coord_equal(expand = TRUE) +
      theme_bw()
  print(p)
  
  # save the plot
  ggsave(paste0("../results/Clusters/TimeCluster_", timestep, "_", i, cluster_variable, ".png"), width = 5, height = 4, dpi = 300)
}
```

```{r}
station_clusters_UHI |>
    ggplot(aes(x = OST_CHTOPO, y = NORD_CHTOPO, fill = station_cluster)) +
      geom_point(      
        color = "black",
        shape = 21,
        size = 4,
        stroke = 0.4) +
      labs(title = "Clustered Stations", 
           fill = "Cluster",
           x = "Longitude", y = "Latitude") +
      theme(legend.position = "bottom") +
      coord_equal(expand = TRUE) +
      scale_fill_brewer(palette = "Dark2") +
      theme_bw()
```

# Leaflet (for daily and hourly)
```{r}
# Base map with layer control
map <- leaflet() |>
  addProviderTiles(providers$Esri.WorldTopoMap, group = "Esri World Topo") |>
  addLayersControl(
    overlayGroups = c(paste("Cluster", 0:3), "Station Clusters"),
    options = layersControlOptions(collapsed = FALSE)
  )

# Add time_cluster layers (Cluster 0 to Cluster 5)
for (i in 0:3) {
  cluster_data <- time_clusters_UHI |>
    filter(time_cluster == i)
  
  pal_uhi <- colorNumeric(palette = viridis(100), domain = c(-1, 5.1), reverse = FALSE)
  
  map <- map |>
    addCircleMarkers(
      data = cluster_data,
      lng = ~OST_CHTOPO,
      lat = ~NORD_CHTOPO,
      fillColor = ~pal_uhi(UHI),
      color = "black",
      fillOpacity = 1,
      stroke = TRUE,
      radius = 6,
      weight = 0.4,
      group = paste("Cluster", i),
      label = ~paste("UHI:", round(UHI, 2))
    )
}

# Add station_clusters_UHI layer
pal_cluster <- colorFactor(palette = "Dark2", domain = station_clusters_UHI$station_cluster)

map <- map |>
  addCircleMarkers(
    data = station_clusters_UHI,
    lng = ~OST_CHTOPO,
    lat = ~NORD_CHTOPO,
    fillColor = ~pal_cluster(station_cluster),
    color = "black",
    fillOpacity = 1,
    stroke = TRUE,
    radius = 6,
    weight = 0.4,
    group = "Station Clusters",
    label = ~paste("Cluster:", station_cluster)
  )

# Add legend for UHI (fixed for all Cluster layers)
map <- map |>
  addLegend(
    position = "bottomright",
    pal = colorNumeric(viridis(100), domain = c(-1, 5.1)),
    values = time_clusters_UHI$UHI,
    title = "UHI [K]",
    group = "UHI Clusters"
  )

# Add legend for Station Clusters
map <- map |>
  addLegend(
    position = "bottomleft",
    pal = pal_cluster,
    values = station_clusters_UHI$station_cluster,
    title = "Station Cluster",
    group = "Station Clusters"
  )

map


```

# distribution of UHI by time cluster
```{r}
# custom color palette for cluster values
if(timesep == "daily"){
  cluster_colors <- c('#a6d854','#fc8d62','#8da0cb','#e78ac3','#66c2a5','#ffd92f')
} else {
  cluster_colors <- c('#fc8d62','#8da0cb','#a6d854','#e78ac3','#66c2a5','#ffd92f')
}

# Look at the distribution of UHI_mean for each time cluster
time_clusters_UHI |>
  ggplot(aes(x = UHI, fill = factor(time_cluster))) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  labs(title = paste("Distribution of UHI by Time Cluster", timestep), fill = "Time Cluster") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = cluster_colors)

ggsave(paste0("../results/Clusters/Distribution_UHI_by_Time_Cluster", timestep, ".png"), width = 5, height = 4, dpi = 300)

```



# Find night that is closest to the cluster means
```{r}
# Compute nightly UHI vectors per station per cluster
nightly_profiles <- all_data_daily_night_clusters |>
  group_by(time_cluster, time, Log_NR) |>
  summarise(UHI = mean(UHI, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = Log_NR, values_from = UHI) |>
  drop_na()  # Remove rows with NA values

library(purrr)

# Get the cluster-wise mean UHI profile (your existing table)
cluster_means <- nightly_profiles |> 
  group_by(time_cluster) |>
  summarise(across(starts_with("Log_"), ~ mean(., na.rm = TRUE)), .groups = "drop")

# str(nightly_profiles)
# str(cluster_means)

# Function to compute Euclidean distance
compute_distance <- function(vec1, vec2) {
  sqrt(sum((vec1 - vec2)^2, na.rm = TRUE))
}

# For each cluster, find the closest night
closest_nights <- nightly_profiles |>
  nest(data = -time_cluster) |>
  left_join(cluster_means |> nest(cluster_mean = -time_cluster), by = "time_cluster") |>
  mutate(
    best_match = map2(data, cluster_mean, function(night_rows, cluster_vec) {
      night_rows |>
        rowwise() |>
        mutate(
          distance = compute_distance(
            c_across(starts_with("Log_")),
            unlist(cluster_vec)
          )
        ) |>
        arrange(distance) |>
        slice(1)
    })
  ) |>
  select(time_cluster, best_match) |>
  unnest(best_match)

closest_nights

closest_nights |>
  group_by(time_cluster) |>
  dplyr::summarize(day = first(time),
            across(starts_with("Log_"), ~ first(.)), .groups = "drop") |>
  # mean over all columns that start with Log_
  pivot_longer(cols = starts_with("Log_"), names_to = "Log_NR", values_to = "UHI") |>
  group_by(time_cluster, day) |>
  dplyr::summarize(UHI_mean = mean(UHI, na.rm = TRUE), .groups = "drop")

```

# visualize the uhi profile differences
```{r}
# Cluster means to long format
cluster_means_long <- cluster_means |>
  pivot_longer(cols = starts_with("Log_"), names_to = "Log_NR", values_to = "UHI_mean")

# Best-matching nights to long format
best_matches_long <- closest_nights |>
  group_by(time_cluster) |>
  summarize(across(starts_with("Log_"), ~ first(.)), .groups = "drop") |>
  pivot_longer(cols = starts_with("Log_"), names_to = "Log_NR", values_to = "UHI_night")

plot_data <- left_join(best_matches_long, cluster_means_long,
                       by = c("time_cluster", "Log_NR"))



# Define your custom cluster colors (you can use only the first 4 if you have 4 clusters)
cluster_colors <- c('#a6d854','#fc8d62','#8da0cb','#e78ac3')
night_colors <- c('#66a61e','#d94f22','#5560a3','#e7339a')
#7570b3
#e7298a
#66a61e)

# Reuse plot_data from earlier step (joined long-format cluster means and best matches)

ggplot(plot_data, aes(x = Log_NR)) +
  # Cluster mean: thick solid line
  geom_line(aes(y = UHI_mean, group = time_cluster, color = time_cluster), linewidth = 1.2, alpha=0.5) +
  # Best-match night: dashed line
  geom_line(aes(y = UHI_night, group = time_cluster, color = time_cluster), 
            linewidth = 0.6) +
  scale_color_manual(values = cluster_colors,
                    labels = c("0" = "rainy\n2024-09-05\n",
                               "1" = "changeable\n2024-08-27\n",
                               "2" = "sunny\n2024-08-29\n",
                               "3" = "windy + dry\n2024-08-14\n")) +
  theme_minimal() +
  labs(
    title = "Mean UHI Profiles per Time Cluster",
    subtitle = "lighter = Cluster Mean | darker = Best-Matching Night",
    x = "Measurement site",
    y = "mean nighttime UHI [K]",
    color = "Time Cluster"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))

ggsave(paste0("../results/Clusters/UHI_Profile_Comparison_", timestep, ".png"), width = 6, height = 4, dpi = 300)

```

# Create TIFFs of the meteorological predictors of the four representative cluster nights
```{r}
# date <- "2024-08-23" # night with highest mean UHI: 3.13 K
# date <- "2024-07-27"# night with average mean UHI: 1.42 K ## not a good day bc of missing data at a logger
# date <- "2024-09-26"# night with lowest mean UHI: -0.97 K
# date <- "2024-08-11" # night with average UHI: 1.54 K
# date <- "2024-08-13" # night with low UHI: 0.69 K

source("../R/map_helpers.R")
#dates = c("2024-08-23", "2024-08-11", "2024-08-13", "2024-09-26") #highest UHI, average UHI, lowest UHI
dates <- closest_nights |>
  group_by(time_cluster) |>
  summarize(day = as.character(first(time)))
dates <- as.vector(dates$day)

for (date in dates) {
  rasters <- create_meteo_tiffs(date = date, overwrite = TRUE)
  print(paste("Created rasters for date:", date))
}
```



