---
title: "Combine the datasets"
author: 'Noémie Wellinger'
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
source("../R/load_packages.R")
load_packages(c("dplyr", "ggplot2", "climaemet", "openair", "lubridate", "tidyverse"))
```



Classify wind
```{r}
MeteoCH_data <- readRDS("../data/IDAWEB/MeteoCH_Thun.rds")
# Classify wind direction
MeteoCH_data <- MeteoCH_data |>
  mutate(wind_dir_10min_class = case_when(
    wind_dir_10min < 22.5 | wind_dir_10min >= 337.5 ~ "N",
    wind_dir_10min >= 22.5 & wind_dir_10min < 67.5 ~ "NE",
    wind_dir_10min >= 67.5 & wind_dir_10min < 112.5 ~ "E",
    wind_dir_10min >= 112.5 & wind_dir_10min < 157.5 ~ "SE",
    wind_dir_10min >= 157.5 & wind_dir_10min < 202.5 ~ "S",
    wind_dir_10min >= 202.5 & wind_dir_10min < 247.5 ~ "SW",
    wind_dir_10min >= 247.5 & wind_dir_10min < 292.5 ~ "W",
    wind_dir_10min >= 292.5 & wind_dir_10min < 337.5 ~ "NW"
  )) |>
  mutate(wind_dir_h_class = case_when(
    wind_dir_h < 22.5 | wind_dir_h >= 337.5 ~ "N",
    wind_dir_h >= 22.5 & wind_dir_h < 67.5 ~ "NE",
    wind_dir_h >= 67.5 & wind_dir_h < 112.5 ~ "E",
    wind_dir_h >= 112.5 & wind_dir_h < 157.5 ~ "SE",
    wind_dir_h >= 157.5 & wind_dir_h < 202.5 ~ "S",
    wind_dir_h >= 202.5 & wind_dir_h < 247.5 ~ "SW",
    wind_dir_h >= 247.5 & wind_dir_h < 292.5 ~ "W",
    wind_dir_h >= 292.5 & wind_dir_h < 337.5 ~ "NW",
  )) |>
  mutate(wind_dir_10min_NS = ifelse(wind_dir_10min > 90 & wind_dir_10min <= 270, "S", "N"),
         wind_dir_h_NS = ifelse(wind_dir_h > 90 & wind_dir_h <= 270, "S", "N"),
         wind_dir_h_WE= ifelse(wind_dir_h > 0 & wind_dir_h <= 180, "E", "W"),
         wind_dir_10min_WE= ifelse(wind_dir_10min > 0 & wind_dir_10min <= 180, "E", "W"))

saveRDS(MeteoCH_data, "../data/IDAWEB/MeteoCH_Thun.rds")
```

# Extract station information from geospatial layers and join to metadata
```{r}
# read metadata
meta <- read.csv("../data/Metadata/metadata_Thun.csv",
                 fileEncoding = "latin1")
# read land use data
tiff_names <- list.files("../data/Geodata/LULCTopo_Gauss/")
print(tiff_names)
tiff_paths <- paste0("../data/Geodata/LULCTopo_Gauss/",tiff_names)
tiffs <- terra::rast(tiff_paths)
# Set layer names based on file names (without extension)
names(tiffs) <- tools::file_path_sans_ext(tiff_names)

spat_points <- meta |> dplyr::select(c(LV_03_E,LV_03_N))
extracted <- terra::extract(tiffs,spat_points)

meta <- meta |>
  dplyr::bind_cols(extracted) |>
  # for all variables that contain BH_NA
  mutate(across(contains("BH_NA"), ~ ifelse(is.na(.), 0, .)))

str(meta)

write.csv(meta, "../data/Metadata/metadata_Thun_geodata.csv")

geospatial_predictors <- meta[, 14:123]
# Save geospatial predictors
write.csv(geospatial_predictors, "../data/final_predictors/geospatial_predictors.csv", row.names = FALSE)
```

Water temperature data
```{r}
# BAFU water temperature
Water_data <- read.csv(
  "../data/external/Water - BAFU/2030_Wassertemperatur_10-Min-Mittel_2024-05-01_2024-09-30.csv",
  skip = 9, # delete metadata
  header = FALSE,
  sep = ";",
  dec = ".",
  fill = TRUE,
  fileEncoding = "latin1",  # Handles ä, ö, ü, ß etc.
  col.names = c("Stationsname", "Stationsnummer", "Parameter", "Zeitreihe",
                "Parametereinheit", "Gewässer", "Zeitstempel",
                "Zeitpunkt_des_Auftretens", "Wert", "Freigabestatus") #There is a blank column in the csv. This code helps to prevent error
)

Water_data_2 <- read.csv(
  "../data/external/Water - BAFU/2030_Wassertemperatur_10-Min-Mittel_2024-10-01_2024-12-01.csv",
  skip = 9, # delete metadata
  header = FALSE,
  sep = ";",
  dec = ".",
  fill = TRUE,
  fileEncoding = "latin1",  # Handles ä, ö, ü, ß etc.
  col.names = c("Stationsname", "Stationsnummer", "Parameter", "Zeitreihe",
                "Parametereinheit", "Gewässer", "Zeitstempel",
                "Zeitpunkt_des_Auftretens", "Wert", "Freigabestatus") #There is a blank column in the csv. This code helps to prevent error
)

# Combine the two datasets
Water_data <- rbind(Water_data, Water_data_2)
Water_data <- Water_data |>
  mutate(time = as.POSIXct(Zeitstempel, format = "%Y-%m-%d %H:%M:%S",  tz = "Europe/Zurich"))
Water_data <- Water_data |>
  select(time, Wert) |>
  rename(water_temp = Wert) |>
  filter(time>= "2024-05-01 00:00:00" & time <= "2024-10-31 23:59:59")

# ALPLAKES mean lake surface temperature

# Temporarily switch to English locale to parse the time format of the following csv correctly
old_locale <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME", "C")

alplakes_data <- read.csv(
  "../data/external/ALPLAKES/ALPLAKES_Average_Lake_Surface_Temperature_LakeThun_Simstrat.csv",
  stringsAsFactors = FALSE,
  fileEncoding = "UTF-8",  # Handles ä, ö, ü, ß etc.
) |>
    mutate(
      water_temp_surface = `Temperature...C.`,
      time_clean = sub(" GMT[+-][0-9]{4}.*$", "", `time...`),  # remove everything in parentheses
      time_posix = as.POSIXct(time_clean, format = "%a %b %d %Y %H:%M:%S", tz = "Europe/Zurich"),
      time = floor_date(time_posix, unit = "hour")) |>
  dplyr::select(time, water_temp_surface)

# Restore original locale
Sys.setlocale("LC_TIME", old_locale)


# Aggregate to hourly data
Water_data_hourly <- Water_data |>
    #mutate(time_hour = floor_date(time, unit = "hour")) |> # round to nearest hour (floor)
  mutate(time = as.POSIXct(time)) |> 
  mutate(time = floor_date(time - minutes(1), unit = "hour") + hours(1)) |>
  group_by(time) |>
  dplyr::summarise(water_temp = mean(water_temp, na.rm = TRUE))

# Merge the two datasets
Water_data_hourly <- Water_data_hourly |>
  left_join(alplakes_data, by = "time")
# interpolate water surface temperature 
Water_data_hourly <- Water_data_hourly |>
   mutate(water_temp_surface = zoo::na.approx(water_temp_surface, x = time, na.rm = FALSE))

visdat::vis_miss(Water_data_hourly)
```

Load the Thun Network data
```{r}
# Thun_Network_data <- read.csv("../data/Thun Messnetz 2024/10min_long_2024-05-01_2024-10-30.csv") |>
#   mutate(time = as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OS", tz="UTC"),
#          timeshift = as.POSIXct(timeshift, format = "%Y-%m-%dT%H:%M:%OS", tz="UTC")) |>
#   mutate(time = with_tz(time, tzone = "Europe/Zurich"),
#          timeshift = with_tz(timeshift, tzone = "Europe/Zurich")) #should start at 2024-05-01 02:00

Thun_Network_data_hourly <- read.csv("../data/Thun Messnetz 2024/Thun_hourly_summary_2024-05-01_2024-10-30.csv",
                                     fileEncoding = "latin1") |>   # Handles ä, ö, ü, ß etc.
  mutate(time_hour = as.POSIXct(time_hour, format = "%Y-%m-%dT%H:%M:%OS", tz="UTC")) |>
  mutate(time_hour = with_tz(time_hour, tzone = "Europe/Zurich")) |> #should start at 2024-05-01 02:00
  rename(time = time_hour)

Thun_Network_data_daily <- read.csv("../data/Thun Messnetz 2024/Thun_daily_summary_2024-05-01_2024-10-30.csv",
                                     fileEncoding = "latin1") |>   # Handles ä, ö, ü, ß etc.
  mutate(time = as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OS", tz="UTC")) |>
  mutate(time = with_tz(time, tzone = "Europe/Zurich")) #should start at 2024-05-01 02:00
```

Load the wind data from Strandbad
```{r}
strandbad <- read.csv("../data/external/Thunerwetter/hourly_wind_24-05-01_24-10-31.CSV") 

# Lookup table for wind direction in degrees
dir_lookup <- c(
  "N" = 0,
  "N-NO" = 22.5,
  "NO" = 45,
  "O-NO" = 67.5,
  "O" = 90,
  "O-SO" = 112.5,
  "SO" = 135,
  "S-SO" = 157.5,
  "S" = 180,
  "S-SW" = 202.5,
  "SW" = 225,
  "W-SW" = 247.5,
  "W" = 270,
  "W-NW" = 292.5,
  "NW" = 315,
  "N-NW" = 337.5
)

strandbad <- strandbad |>
  mutate(
    time = dmy_hm(paste(Datum, Zeit), tz = "Europe/Zurich"),
    wind_dir_lake = dir_lookup[wind_dir],
    wind_vel_lake = wind_vel
  ) |>
  select(time, wind_dir_lake, wind_vel_lake)
```

# Aggregate MeteoCH data and add memory/rolling means
```{r}
MeteoCH_data <- readRDS("../data/IDAWEB/MeteoCH_Thun.rds")
print(tz(MeteoCH_data$time))

tz(MeteoCH_data$time) <- "UTC"  # not mutate, direct assignment!
print(tz(MeteoCH_data$time))

MeteoCH_data <- MeteoCH_data |>
  mutate(time = with_tz(time, tzone = "Europe/Zurich")) |> #should start at 2024-05-01 02:00
  select(-stn)


# MeteoCH_data_10min <- MeteoCH_data |>
#   select(time, temp_air_2m_10min, temp_pot_2m_10m, precip_10min, RH_2m_10min, glob_rad_10min, airpress_10min, wind_vel_10min_ms, wind_vel_10min_kmh, wind_dir_10min, wind_dir_10min_class, wind_dir_10min_NS, CAP9, CAP18, CAP27, GWT8_Z500, GWT10_SLP, GWT10_Z500, GWT18_SLP, GWT18_Z500, GWT26_SLP, GWT26_Z500) |>
#   drop_na()

MeteoCH_data_hourly <- MeteoCH_data |>
  select(time, temp_air_2m_h, precip_last1h, RH_2m_h, glob_rad_h, airpress_h, wind_vel_h_kmh, wind_dir_h, wind_dir_h_class, wind_dir_h_NS, wind_dir_h_WE, CAP9, CAP18, CAP27, GWT8_Z500, GWT10_SLP, GWT10_Z500, GWT18_SLP, GWT18_Z500, GWT26_SLP, GWT26_Z500) |>
  drop_na() |> # drop rows with 10 min values
  rename(air_temp = temp_air_2m_h,
         pressure = airpress_h,
         wind_vel = wind_vel_h_kmh,
         wind_dir = wind_dir_h,
         rad = glob_rad_h,
         humidity = RH_2m_h,
         precip = precip_last1h,
         wind_dir_class = wind_dir_h_class,
         wind_dir_NS = wind_dir_h_NS,
         wind_dir_WE = wind_dir_h_WE) |>
  # join with water_temp data
  left_join(Water_data_hourly, by = "time") |>
  left_join(strandbad, by = "time") |>
    mutate(CAP9 = as.factor(CAP9),
         CAP18 = as.factor(CAP18),
         CAP27 = as.factor(CAP27),
         GWT8_Z500 = as.factor(GWT8_Z500),
         GWT10_SLP = as.factor(GWT10_SLP),
         GWT10_Z500 = as.factor(GWT10_Z500),
         GWT18_SLP = as.factor(GWT18_SLP),
         GWT18_Z500 = as.factor(GWT18_Z500),
         GWT26_SLP = as.factor(GWT26_SLP),
         GWT26_Z500 = as.factor(GWT26_Z500)
    )

 MeteoCH_data_hourly <- MeteoCH_data_hourly |>
    arrange(time)|> # make sure it's in the right order
    mutate(
      air_temp_8h = zoo::rollmean(air_temp, k = 8, fill = NA, align = "right"),
      air_temp_16h = zoo::rollmean(air_temp, k = 16, fill = NA, align = "right"),
      air_temp_1d = zoo::rollmean(air_temp, k = 24 * 1, fill = NA, align = "right"),
      air_temp_3d = zoo::rollmean(air_temp, k = 24 * 3, fill = NA, align = "right"),
      pressure_8h = zoo::rollmean(pressure, k = 8, fill = NA, align = "right"),
      pressure_16h = zoo::rollmean(pressure, k = 16, fill = NA, align = "right"),
      pressure_1d = zoo::rollmean(pressure, k = 24 * 1, fill = NA, align = "right"),
      pressure_3d = zoo::rollmean(pressure, k = 24 * 3, fill = NA, align = "right"),
      wind_vel_8h = zoo::rollmean(wind_vel, k = 8, fill = NA, align = "right"),
      wind_vel_16h = zoo::rollmean(wind_vel, k = 16, fill = NA, align = "right"),
      wind_vel_1d = zoo::rollmean(wind_vel, k = 24 * 1, fill = NA, align = "right"),
      wind_vel_3d = zoo::rollmean(wind_vel, k = 24 * 3, fill = NA, align = "right"),
      wind_vel_lake_8h = zoo::rollmean(wind_vel_lake, k = 8, fill = NA, align = "right"),
      wind_vel_lake_16h = zoo::rollmean(wind_vel_lake, k = 16, fill = NA, align = "right"),
      wind_vel_lake_1d = zoo::rollmean(wind_vel_lake, k = 24 * 1, fill = NA, align = "right"),
      wind_vel_lake_3d = zoo::rollmean(wind_vel_lake, k = 24 * 3, fill = NA, align = "right"),
      rad_8h = zoo::rollmean(rad, k = 8, fill = NA, align = "right"),
      rad_16h = zoo::rollmean(rad, k = 16, fill = NA, align = "right"),
      rad_1d = zoo::rollmean(rad, k = 24 * 1, fill = NA, align = "right"),
      rad_3d = zoo::rollmean(rad, k = 24 * 3, fill = NA, align = "right"),
      humidity_8h = zoo::rollmean(humidity, k = 8, fill = NA, align = "right"),
      humidity_16h = zoo::rollmean(humidity, k = 16, fill = NA, align = "right"),
      humidity_1d = zoo::rollmean(humidity, k = 24 * 1, fill = NA, align = "right"),
      humidity_3d = zoo::rollmean(humidity, k = 24 * 3, fill = NA, align = "right"),
      precip_8h = zoo::rollsum(precip, k = 8, fill = NA, align = "right"),
      precip_16h = zoo::rollsum(precip, k = 16, fill = NA, align = "right"),
      precip_1d = zoo::rollsum(precip, k = 24 * 1, fill = NA, align = "right"),
      precip_3d = zoo::rollsum(precip, k = 24 * 3, fill = NA, align = "right"),
      water_temp_8h = zoo::rollmean(water_temp, k = 8, fill = NA, align = "right"),
      water_temp_16h = zoo::rollmean(water_temp, k = 16, fill = NA, align = "right"),
      water_temp_1d = zoo::rollmean(water_temp, k = 24 * 1, fill = NA, align = "right"),
      water_temp_3d = zoo::rollmean(water_temp, k = 24 * 3, fill = NA, align = "right"),
      water_temp_surface_8h = zoo::rollmean(water_temp_surface, k = 8, fill = NA, align = "right"),
      water_temp_surface_16h = zoo::rollmean(water_temp_surface, k = 16, fill = NA, align = "right"),
      water_temp_surface_1d = zoo::rollmean(water_temp_surface, k = 24 * 1, fill = NA, align = "right"),
      water_temp_surface_3d = zoo::rollmean(water_temp_surface, k = 24 * 3, fill = NA, align = "right")) |>
    mutate_at(vars(starts_with("precip")), ~ ifelse(. < 0.1, 0, .))
 
source("../R/wind_helpers.R") 
MeteoCH_data_hourly <- roll_mean_wind_dir(
  df = MeteoCH_data_hourly,
  dir_col = "wind_dir",
  speed_col = "wind_vel",
  k = c(8, 16, 24, 72),
  suffix = c("8h", "16h", "1d", "3d")
)
MeteoCH_data_hourly <- roll_mean_wind_dir(
  df = MeteoCH_data_hourly,
  dir_col = "wind_dir_lake",
  speed_col = "wind_vel_lake",
  k = c(8, 16, 24, 72),
  suffix = c("8h", "16h", "1d", "3d")
) 
```

# Aggregate the meteodata to daily nighttime data
```{r}
MeteoCH_data_night <- MeteoCH_data_hourly |>
  filter(hour(time) > 22 | hour(time) <= 6) |>
  # shift time 7h back
  mutate(time = time - hours(7)) |>
  group_by(date = as.Date(time)) |>
  select(-air_temp, -precip, -humidity, -rad, -pressure, 
         -wind_vel, -wind_dir, -water_temp, -water_temp_surface, -wind_dir_lake, -wind_vel_lake, -wind_dir_class, -wind_dir_NS, -wind_dir_WE,
         -air_temp_16h, -precip_16h, -humidity_16h, -rad_16h, -pressure_16h, 
         -wind_vel_16h, -wind_dir_16h, -water_temp_16h, -water_temp_surface_16h, -wind_vel_lake_16h, -wind_dir_lake_16h) |>
  # summarize all variables by taking the value of the last hour (i.e. hour(time = 23))
  slice_max(order_by = time, n = 1) |>
  ungroup() |>
  # add a column that marks wether it's day or night
  mutate(daynight = as.factor("n"))


# Classify precipitation
MeteoCH_data_night <- MeteoCH_data_night |>
  mutate(precip_8h_boolean = ifelse(precip_8h > 0.3, 1, 0),
         precip_1d_boolean = ifelse(precip_1d > 0.3, 1, 0),
         precip_3d_boolean = ifelse(precip_3d > 0.3, 1, 0))

# Classify Wind
MeteoCH_data_night <- MeteoCH_data_night |>
  mutate(
    wind_dir_8h_class = classify_wind_dir(wind_dir_8h),
    wind_dir_8h_NS = ifelse(wind_dir_8h > 90 & wind_dir_8h <= 270, "S", "N"),
    wind_dir_8h_WE = ifelse(wind_dir_8h > 0 & wind_dir_8h <= 180, "E", "W"),
    wind_S_8h = ifelse(wind_dir_8h_NS == "S", wind_vel_8h, 0),
    wind_N_8h = ifelse(wind_dir_8h_NS == "N", wind_vel_8h, 0),
    wind_W_8h = ifelse(wind_dir_8h_WE == "W", wind_vel_8h, 0),
    wind_E_8h = ifelse(wind_dir_8h_WE == "E", wind_vel_8h, 0),
    wind_dir_1d_class = classify_wind_dir(wind_dir_1d),
    wind_dir_1d_NS = ifelse(wind_dir_1d > 90 & wind_dir_1d <= 270, "S", "N"),
    wind_dir_1d_WE = ifelse(wind_dir_1d > 0 & wind_dir_1d <= 180, "E", "W"),
    wind_S_1d = ifelse(wind_dir_1d_NS == "S", wind_vel_1d, 0),
    wind_N_1d = ifelse(wind_dir_1d_NS == "N", wind_vel_1d, 0),
    wind_W_1d = ifelse(wind_dir_1d_WE == "W", wind_vel_1d, 0),
    wind_E_1d = ifelse(wind_dir_1d_WE == "E", wind_vel_1d, 0),
    wind_dir_3d_class = classify_wind_dir(wind_dir_3d),
    wind_dir_3d_NS = ifelse(wind_dir_3d > 90 & wind_dir_3d <= 270, "S", "N"),
    wind_dir_3d_WE = ifelse(wind_dir_3d > 0 & wind_dir_3d <= 180, "E", "W"),
    wind_S_3d = ifelse(wind_dir_3d_NS == "S", wind_vel_3d, 0),
    wind_N_3d = ifelse(wind_dir_3d_NS == "N", wind_vel_3d, 0),
    wind_W_3d = ifelse(wind_dir_3d_WE == "W", wind_vel_3d, 0),
    wind_E_3d = ifelse(wind_dir_3d_WE == "E", wind_vel_3d, 0)
    ) |>
  # Classify lake wind
  mutate(
    wind_dir_lake_8h_class = classify_wind_dir(wind_dir_lake_8h),
    wind_dir_lake_8h_NS = ifelse(wind_dir_lake_8h > 90 & wind_dir_lake_8h <= 270, "S", "N"),
    wind_dir_lake_8h_WE = ifelse(wind_dir_lake_8h > 0 & wind_dir_lake_8h <= 180, "E", "W"),
    wind_S_lake_8h = ifelse(wind_dir_lake_8h_NS == "S", wind_vel_lake_8h, 0),
    wind_N_lake_8h = ifelse(wind_dir_lake_8h_NS == "N", wind_vel_lake_8h, 0),
    wind_W_lake_8h = ifelse(wind_dir_lake_8h_WE == "W", wind_vel_lake_8h, 0),
    wind_E_lake_8h = ifelse(wind_dir_lake_8h_WE == "E", wind_vel_lake_8h, 0),
    wind_dir_lake_1d_class = classify_wind_dir(wind_dir_lake_1d),
    wind_dir_lake_1d_NS = ifelse(wind_dir_lake_1d > 90 & wind_dir_lake_1d <= 270, "S", "N"),
    wind_dir_lake_1d_WE = ifelse(wind_dir_lake_1d > 0 & wind_dir_lake_1d <= 180, "E", "W"),
    wind_S_lake_1d = ifelse(wind_dir_lake_1d_NS == "S", wind_vel_lake_1d, 0),
    wind_N_lake_1d = ifelse(wind_dir_lake_1d_NS == "N", wind_vel_lake_1d, 0),
    wind_W_lake_1d = ifelse(wind_dir_lake_1d_WE == "W", wind_vel_lake_1d, 0),
    wind_E_lake_1d = ifelse(wind_dir_lake_1d_WE == "E", wind_vel_lake_1d, 0),
    wind_dir_lake_3d_class = classify_wind_dir(wind_dir_lake_3d),
    wind_dir_lake_3d_NS = ifelse(wind_dir_lake_3d > 90 & wind_dir_lake_3d <= 270, "S", "N"),
    wind_dir_lake_3d_WE = ifelse(wind_dir_lake_3d > 0 & wind_dir_lake_3d <= 180, "E", "W"),
    wind_S_lake_3d = ifelse(wind_dir_lake_3d_NS == "S", wind_vel_lake_3d, 0),
    wind_N_lake_3d = ifelse(wind_dir_lake_3d_NS == "N", wind_vel_lake_3d, 0),
    wind_W_lake_3d = ifelse(wind_dir_lake_3d_WE == "W", wind_vel_lake_3d, 0),
    wind_E_lake_3d = ifelse(wind_dir_lake_3d_WE == "E", wind_vel_lake_3d, 0)
  ) 
```

# Aggregate the meteodata to daily daytime data
```{r}
MeteoCH_data_day <- MeteoCH_data_hourly |>
  filter(hour(time) > 6 & hour(time) <= 22) |>
  group_by(date = as.Date(time)) |>
  select(-air_temp, -precip, -humidity, -rad, -pressure, 
         -wind_vel, -wind_dir, -water_temp, -water_temp_surface, -wind_dir_lake, -wind_vel_lake, -wind_dir_class, -wind_dir_NS, -wind_dir_WE,
         -air_temp_8h, -precip_8h, -humidity_8h, -rad_8h, -pressure_8h, 
         -wind_vel_8h, -wind_dir_8h, -water_temp_8h, -water_temp_surface_8h, -wind_dir_lake_8h, -wind_vel_lake_8h) |>
  # summarize all variables by taking the value of the last hour (i.e. hour(time = 22))
  slice_max(order_by = time, n = 1) |>
  ungroup() |>
  mutate(daynight = as.factor("d")) |>
  # Classify precipitation
  mutate(
    precip_16h_boolean = ifelse(precip_16h > 0.3, 1, 0),
    precip_1d_boolean = ifelse(precip_1d > 0.3, 1, 0),
    precip_3d_boolean = ifelse(precip_3d > 0.3, 1, 0)
  ) |>
  # Classify wind
  mutate(
    wind_dir_16h_class = classify_wind_dir(wind_dir_16h),
    wind_dir_16h_NS = ifelse(wind_dir_16h > 90 & wind_dir_16h <= 270, "S", "N"),
    wind_dir_16h_WE = ifelse(wind_dir_16h > 0 & wind_dir_16h <= 180, "E", "W"),
    wind_S_16h = ifelse(wind_dir_16h_NS == "S", wind_vel_16h, 0),
    wind_N_16h = ifelse(wind_dir_16h_NS == "N", wind_vel_16h, 0),
    wind_W_16h = ifelse(wind_dir_16h_WE == "W", wind_vel_16h, 0),
    wind_E_16h = ifelse(wind_dir_16h_WE == "E", wind_vel_16h, 0),
    wind_dir_1d_class = classify_wind_dir(wind_dir_1d),
    wind_dir_1d_NS = ifelse(wind_dir_1d > 90 & wind_dir_1d <= 270, "S", "N"),
    wind_dir_1d_WE = ifelse(wind_dir_1d > 0 & wind_dir_1d <= 180, "E", "W"),
    wind_S_1d = ifelse(wind_dir_1d_NS == "S", wind_vel_1d, 0),
    wind_N_1d = ifelse(wind_dir_1d_NS == "N", wind_vel_1d, 0),
    wind_W_1d = ifelse(wind_dir_1d_WE == "W", wind_vel_1d, 0),
    wind_E_1d = ifelse(wind_dir_1d_WE == "E", wind_vel_1d, 0),
    wind_dir_3d_class = classify_wind_dir(wind_dir_3d),
    wind_dir_3d_NS = ifelse(wind_dir_3d > 90 & wind_dir_3d <= 270, "S", "N"),
    wind_dir_3d_WE = ifelse(wind_dir_3d > 0 & wind_dir_3d <= 180, "E", "W"),
    wind_S_3d = ifelse(wind_dir_3d_NS == "S", wind_vel_3d, 0),
    wind_N_3d = ifelse(wind_dir_3d_NS == "N", wind_vel_3d, 0),
    wind_W_3d = ifelse(wind_dir_3d_WE == "W", wind_vel_3d, 0),
    wind_E_3d = ifelse(wind_dir_3d_WE == "E", wind_vel_3d, 0)
  ) |>
  # Classify lake wind
  mutate(
    wind_dir_lake_16h_class = classify_wind_dir(wind_dir_lake_16h),
    wind_dir_lake_16h_NS = ifelse(wind_dir_lake_16h > 90 & wind_dir_lake_16h <= 270, "S", "N"),
    wind_dir_lake_16h_WE = ifelse(wind_dir_lake_16h > 0 & wind_dir_lake_16h <= 180, "E", "W"),
    wind_S_lake_16h = ifelse(wind_dir_lake_16h_NS == "S", wind_vel_lake_16h, 0),
    wind_N_lake_16h = ifelse(wind_dir_lake_16h_NS == "N", wind_vel_lake_16h, 0),
    wind_W_lake_16h = ifelse(wind_dir_lake_16h_WE == "W", wind_vel_lake_16h, 0),
    wind_E_lake_16h = ifelse(wind_dir_lake_16h_WE == "E", wind_vel_lake_16h, 0),
    wind_dir_lake_1d_class = classify_wind_dir(wind_dir_lake_1d),
    wind_dir_lake_1d_NS = ifelse(wind_dir_lake_1d > 90 & wind_dir_lake_1d <= 270, "S", "N"),
    wind_dir_lake_1d_WE = ifelse(wind_dir_lake_1d > 0 & wind_dir_lake_1d <= 180, "E", "W"),
    wind_S_lake_1d = ifelse(wind_dir_lake_1d_NS == "S", wind_vel_lake_1d, 0),
    wind_N_lake_1d = ifelse(wind_dir_lake_1d_NS == "N", wind_vel_lake_1d, 0),
    wind_W_lake_1d = ifelse(wind_dir_lake_1d_WE == "W", wind_vel_lake_1d, 0),
    wind_E_lake_1d = ifelse(wind_dir_lake_1d_WE == "E", wind_vel_lake_1d, 0),
    wind_dir_lake_3d_class = classify_wind_dir(wind_dir_lake_3d),
    wind_dir_lake_3d_NS = ifelse(wind_dir_lake_3d > 90 & wind_dir_lake_3d <= 270, "S", "N"),
    wind_dir_lake_3d_WE = ifelse(wind_dir_lake_3d > 0 & wind_dir_lake_3d <= 180, "E", "W"),
    wind_S_lake_3d = ifelse(wind_dir_lake_3d_NS == "S", wind_vel_lake_3d, 0),
    wind_N_lake_3d = ifelse(wind_dir_lake_3d_NS == "N", wind_vel_lake_3d, 0),
    wind_W_lake_3d = ifelse(wind_dir_lake_3d_WE == "W", wind_vel_lake_3d, 0),
    wind_E_lake_3d = ifelse(wind_dir_lake_3d_WE == "E", wind_vel_lake_3d, 0)
  )

```


# Join the hourly Thun network data with the other Meteo data
```{r}
# Save the meteo predictors
# all variables of type char as factors
MeteoCH_data_hourly <- MeteoCH_data_hourly |>
  mutate_if(is.character, as.factor)

write.csv(MeteoCH_data_hourly, "../data/final_predictors/Meteo_predictors_hourly.csv")
saveRDS(MeteoCH_data_hourly, "../data/final_predictors/Meteo_predictors_hourly.rds")

#extract only date from time
all_data_hourly <- Thun_Network_data_hourly |>
  select(time, Log_NR, T_mean) |>
  rename(air_temp_LCD = T_mean) |>
  left_join(meta, by = "Log_NR") |>
  left_join(MeteoCH_data_hourly, by = "time") |>
  select(-X, -HuM, -ID) |>
  mutate(UHI = air_temp_LCD - air_temp) |>
  group_by(time) |>
  mutate(UHI_LCD = if (any(Log_NR == "Log_401")) {
            air_temp_LCD - air_temp_LCD[Log_NR == "Log_401"]
          } else {
            NA
          }) |>
  ungroup()

#visdat::vis_miss(all_data_hourly, warn_large_data = FALSE)

write.csv(all_data_hourly, "../data/final_predictors/all_data_hourly.csv")
saveRDS(all_data_hourly, "../data/final_predictors/all_data_hourly.rds")
```

# Join the half-daily data
```{r}
# Clean column names in both data frames
MeteoCH_data_day <- MeteoCH_data_day |>
  rename_with(~ str_replace_all(., "_16h", ""))

MeteoCH_data_night <- MeteoCH_data_night |>
  rename_with(~ str_replace_all(., "_8h", ""))

MeteoCH_data_daily <- MeteoCH_data_day |>
  add_row(MeteoCH_data_night) |>
  arrange(time) |>
  select( -date) |>
  mutate(time = as.Date(time)) |>
# all variables of type char as factors
  mutate_if(is.character, as.factor)

# rename old predictors to match shorter names
name_table <- read.csv("../data/final_predictors/predictors_oldnames_mapping.csv", stringsAsFactors = FALSE)

MeteoCH_data_daily <- MeteoCH_data_daily |>
  rename_with(~ ifelse(
    is.na(match(., name_table$name_old)),
    .,  # keep original name
    name_table$name_new[match(., name_table$name_old)]
  ))

write.csv(MeteoCH_data_daily, "../data/final_predictors/Meteo_predictors_daily.csv")
saveRDS(MeteoCH_data_daily, "../data/final_predictors/Meteo_predictors_daily.rds")

all_data_daily <- Thun_Network_data_daily |>
  select(time, Log_NR, T_mean_day, T_mean_night, UHI_day, UHI_night) |>
  rename(air_temp_LCD_night = T_mean_night,
         air_temp_LCD_day = T_mean_day,
         UHI_LCD_night = UHI_night,
         UHI_LCD_day = UHI_day) |>
  left_join(meta, by = "Log_NR") |>
  
  left_join(MeteoCH_data_daily, by = "time") |>
  select(-X, -HuM, -ID) |>
  # UHI between reference temperature measured by SwissMetNet station and LCD
  mutate(air_temp_LCD = ifelse(daynight == "d", air_temp_LCD_day, air_temp_LCD_night), 
         UHI_LCD = ifelse(daynight == "d", UHI_LCD_day, UHI_LCD_night)) |>
  mutate(UHI_day = air_temp_LCD_day - air_temp,
         UHI_night = air_temp_LCD_night - air_temp,
         UHI = ifelse(daynight == "d", UHI_day, UHI_night)) |>
  select(-UHI_day, -UHI_night, -air_temp_LCD_day, -air_temp_LCD_night, -UHI_LCD_day, -UHI_LCD_night)

visdat::vis_miss(all_data_daily, warn_large_data = FALSE)

all_data_daily <- all_data_daily |>
  rename_with(~ ifelse(
    is.na(match(., name_table$name_old)),
    .,  # keep original name
    name_table$name_new[match(., name_table$name_old)]
  ))


write.csv(all_data_daily, "../data/final_predictors/all_data_daily.csv")
saveRDS(all_data_daily, "../data/final_predictors/all_data_daily.rds")
```


Remove temporary tables *(only when the data is saved and preprocessing finished!)*
```{r}
rm(strandbad, Water_data, Water_data_hourly, 
   MeteoCH_data, MeteoCH_data_hourly, MeteoCH_data_night, MeteoCH_data_day,
   Thun_Network_data_hourly, Thun_Network_data_daily,
   Thun_Network_data, alplakes_data, extracted)
```


