---
title: "Process_IDAWEB_data"
author: 'Noémie Wellinger'
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
source("../R/load_packages.R")
load_packages(c("dplyr", "ggplot2", "climaemet", "openair", "lubridate", "tidyverse"))
```



Classify wind and precipitation
```{r}
MeteoCH_data <- read.csv("../data/IDAWEB/MeteoCH_Thun.csv")
# Convert the time column to a proper datetime format
MeteoCH_data$time <- as.POSIXct(MeteoCH_data$time, format="%Y-%m-%d %H:%M:%S", tz = "UTC")
# Classify wind direction
MeteoCH_data <- MeteoCH_data |>
  mutate(wind_dir_10min_class = case_when(
    wind_dir_10min < 22.5 & wind_dir_10min >= 337.5 ~ "N",
    wind_dir_10min >= 22.5 & wind_dir_10min < 67.5 ~ "NE",
    wind_dir_10min >= 67.5 & wind_dir_10min < 112.5 ~ "E",
    wind_dir_10min >= 112.5 & wind_dir_10min < 157.5 ~ "SE",
    wind_dir_10min >= 157.5 & wind_dir_10min < 202.5 ~ "S",
    wind_dir_10min >= 202.5 & wind_dir_10min < 247.5 ~ "SW",
    wind_dir_10min >= 247.5 & wind_dir_10min < 292.5 ~ "W",
    wind_dir_10min >= 292.5 & wind_dir_10min < 337.5 ~ "NW"
  )) |>
  mutate(wind_dir_h_class = case_when(
    wind_dir_h < 22.5 & wind_dir_h >= 337.5 ~ "N",
    wind_dir_h >= 22.5 & wind_dir_h < 67.5 ~ "NE",
    wind_dir_h >= 67.5 & wind_dir_h < 112.5 ~ "E",
    wind_dir_h >= 112.5 & wind_dir_h < 157.5 ~ "SE",
    wind_dir_h >= 157.5 & wind_dir_h < 202.5 ~ "S",
    wind_dir_h >= 202.5 & wind_dir_h < 247.5 ~ "SW",
    wind_dir_h >= 247.5 & wind_dir_h < 292.5 ~ "W",
    wind_dir_h >= 292.5 & wind_dir_h < 337.5 ~ "NW",
  )) |>
  mutate(wind_dir_d_class = case_when(
    wind_dir_d < 22.5 & wind_dir_d >= 337.5 ~ "N",
    wind_dir_d >= 22.5 & wind_dir_d < 67.5 ~ "NE",
    wind_dir_d >= 67.5 & wind_dir_d < 112.5 ~ "E",
    wind_dir_d >= 112.5 & wind_dir_d < 157.5 ~ "SE",
    wind_dir_d >= 157.5 & wind_dir_d < 202.5 ~ "S",
    wind_dir_d >= 202.5 & wind_dir_d < 247.5 ~ "SW",
    wind_dir_d >= 247.5 & wind_dir_d < 292.5 ~ "W",
    wind_dir_d >= 292.5 & wind_dir_d < 337.5 ~ "NW",
  )) |>
  mutate(wind_dir_10min_NS = ifelse(wind_dir_10min > 90 & wind_dir_10min <= 270, "S", "N"),
         wind_dir_h_NS = ifelse(wind_dir_h > 90 & wind_dir_h <= 270, "S", "N"),
         wind_dir_d_NS = ifelse(wind_dir_d > 90 & wind_dir_d <= 270, "S", "N"))

write.csv(MeteoCH_data, "../data/IDAWEB/MeteoCH_Thun.csv")

```

# Extract station information from geospatial layers and join to metadata
```{r}
# read metadata
meta <- read.csv("../data/Metadata/metadata_Thun.csv")

# read land use data
tiff_names <- list.files("../data/Geodata/LULCTopo_Thun/")
print(tiff_names)
tiff_paths <- paste0("../data/Geodata/LULCTopo_Thun/",tiff_names)
tiffs <- terra::rast(tiff_paths)
# Set layer names based on file names (without extension)
names(tiffs) <- tools::file_path_sans_ext(tiff_names)

spat_points <- meta |> dplyr::select(c(LV_03_E,LV_03_N))
extracted <- terra::extract(tiffs,spat_points)

meta <- meta |>
  dplyr::bind_cols(extracted)

str(meta)

write.csv(meta, "../data/Metadata/metadata_Thun_geodata.csv")
```

Water temperature data
```{r}
# BAFU water temperature
Water_data <- read.csv(
  "../data/external/Water - BAFU/2030_Wassertemperatur_10-Min-Mittel_2024-05-01_2024-09-30.csv",
  skip = 9, # delete metadata
  header = FALSE,
  sep = ";",
  dec = ".",
  fill = TRUE,
  fileEncoding = "latin1",  # Handles ä, ö, ü, ß etc.
  col.names = c("Stationsname", "Stationsnummer", "Parameter", "Zeitreihe",
                "Parametereinheit", "Gewässer", "Zeitstempel",
                "Zeitpunkt_des_Auftretens", "Wert", "Freigabestatus") #There is a blank column in the csv. This code helps to prevent error
)
Water_data <- Water_data |>
  mutate(time = as.POSIXct(Water_data$Zeitstempel, format = "%Y-%m-%d %H:%M:%S",  tz = "Europe/Zurich")) |>
  select(time, water_temp = Wert)



# ALPLAKES mean lake surface temperature

# Temporarily switch to English locale to parse the time format of the following csv correctly
old_locale <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME", "C")

alplakes_data <- read.csv(
  "../data/external/ALPLAKES/ALPLAKES_Average_Lake_Surface_Temperature_LakeThun_Simstrat.csv",
  stringsAsFactors = FALSE,
  fileEncoding = "UTF-8",  # Handles ä, ö, ü, ß etc.
) |>
    mutate(
      water_temp_surface = `Temperature...C.`,
      time_clean = sub(" GMT[+-][0-9]{4}.*$", "", `time...`),  # remove everything in parentheses
      time_posix = as.POSIXct(time_clean, format = "%a %b %d %Y %H:%M:%S", tz = "Europe/Zurich"),
      time = floor_date(time_posix, unit = "hour")) |>
  dplyr::select(time, water_temp_surface)

# Restore original locale
Sys.setlocale("LC_TIME", old_locale)


# Aggregate to hourly data
Water_data_hourly <- Water_data |>
    #mutate(time_hour = floor_date(time, unit = "hour")) |> # round to nearest hour (floor)
  mutate(time = floor_date(time - minutes(1), unit = "hour") + hours(1)) |>
  group_by(time) |>
  summarise(water_temp = mean(water_temp, na.rm = TRUE))

# Merge the two datasets
Water_data_hourly <- Water_data_hourly |>
  left_join(alplakes_data, by = "time")
# interpolate water surface temperature 
Water_data_hourly <- Water_data_hourly |>
   mutate(water_temp_surface = zoo::na.approx(water_temp_surface, x = time, na.rm = FALSE))

# Water_data_daily_night <- Water_data_hourly |>
#   filter(hour(time) >= 22 | hour(time) <= 6) |>
#   # shift time 7h back
#   mutate(time = time - hours(7)) |>
#   group_by(time = as.Date(time)) |>
#   summarise(water_temp = mean(water_temp, na.rm = TRUE),
#             water_temp_surface = mean(water_temp_surface, na.rm = TRUE)) |>
#   ungroup()
```

Load the Thun Network data
```{r}
Thun_Network_data <- read.csv("../data/Thun Messnetz 2024/10min_long_2024-05-01_2024-10-30.csv") |>
  mutate(time = as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OS", tz="UTC"),
         timeshift = as.POSIXct(timeshift, format = "%Y-%m-%dT%H:%M:%OS", tz="UTC")) |>
  mutate(time = with_tz(time, tzone = "Europe/Zurich"),
         timeshift = with_tz(timeshift, tzone = "Europe/Zurich")) #should start at 2024-05-01 02:00

Thun_Network_data_hourly <- read.csv("../data/Thun Messnetz 2024/Thun_hourly_summary_2024-05-01_2024-10-30.csv") |>
  mutate(time_hour = as.POSIXct(time_hour, format = "%Y-%m-%dT%H:%M:%OS", tz="UTC")) |>
  mutate(time_hour = with_tz(time_hour, tzone = "Europe/Zurich")) |> #should start at 2024-05-01 02:00
  rename(time = time_hour)

Thun_Network_data_daily <- read.csv("../data/Thun Messnetz 2024/Thun_daily_summary_2024-05-01_2024-10-30.csv") |>
  mutate(time_day = as.POSIXct(time_day, format = "%Y-%m-%dT%H:%M:%OS", tz="UTC")) |>
  mutate(time_day = with_tz(time_day, tzone = "Europe/Zurich")) |> #should start at 2024-05-01 02:00
  rename(time = time_day)
```

# Aggregate MeteoCH data and add memory/rolling means
```{r}
MeteoCH_data <- read.csv("../data/IDAWEB/MeteoCH_Thun.csv") |>
  mutate(time = as.POSIXct(time, format="%d.%m.%Y %H:%M", tz = "UTC")) |> 
  mutate(time = with_tz(time, tzone = "Europe/Zurich")) |> #should start at 2024-05-01 02:00
  select(-X.1, -X, -stn)

MeteoCH_data_10min <- MeteoCH_data |>
  select(time, temp_air_2m_10min, temp_pot_2m_10m, precip_10min, RH_2m_10min, glob_rad_10min, airpress_10min, wind_vel_10min_ms, wind_vel_10min_kmh, wind_dir_10min, wind_dir_10min_class, wind_dir_10min_NS, CAP9, CAP18, CAP27, GWT8_Z500, GWT10_SLP, GWT10_Z500, GWT18_SLP, GWT18_Z500, GWT26_SLP, GWT26_Z500) |>
  drop_na()

MeteoCH_data_hourly <- MeteoCH_data |>
  select(time, temp_air_2m_h, precip_last1h, precip_last24h, precip_last72h, RH_2m_h, glob_rad_h, airpress_h, wind_vel_h_kmh, wind_dir_h, wind_dir_h_class, wind_dir_h_NS, CAP9, CAP18, CAP27, GWT8_Z500, GWT10_SLP, GWT10_Z500, GWT18_SLP, GWT18_Z500, GWT26_SLP, GWT26_Z500) |>
  drop_na() |>
  rename(air_temp = temp_air_2m_h,
         pressure = airpress_h,
         wind_vel = wind_vel_h_kmh,
         wind_dir = wind_dir_h,
         rad = glob_rad_h,
         humidity = RH_2m_h,
         precip = precip_last1h) |>
  # join with water_temp data
  left_join(Water_data_hourly, by = "time")

 MeteoCH_data_hourly <- MeteoCH_data_hourly |>
    arrange(time)|> # make sure it's in the right order
    mutate(
      mean_air_temp_6h = zoo::rollmean(air_temp, k = 6, fill = NA, align = "right"),
      mean_air_temp_12h = zoo::rollmean(air_temp, k = 12, fill = NA, align = "right"),
      mean_air_temp_1d = zoo::rollmean(air_temp, k = 24 * 1, fill = NA, align = "right"),
      mean_air_temp_3d = zoo::rollmean(air_temp, k = 24 * 3, fill = NA, align = "right"),
      mean_air_temp_5d = zoo::rollmean(air_temp, k = 24 * 5, fill = NA, align = "right"),
      mean_pressure_6h = zoo::rollmean(pressure, k = 6, fill = NA, align = "right"),
      mean_pressure_12h = zoo::rollmean(pressure, k = 12, fill = NA, align = "right"),
      mean_pressure_1d = zoo::rollmean(pressure, k = 24 * 1, fill = NA, align = "right"),
      mean_pressure_3d = zoo::rollmean(pressure, k = 24 * 3, fill = NA, align = "right"),
      mean_pressure_5d = zoo::rollmean(pressure, k = 24 * 5, fill = NA, align = "right"),
      mean_wind_vel_6h = zoo::rollmean(wind_vel, k = 6, fill = NA, align = "right"),
      mean_wind_vel_12h = zoo::rollmean(wind_vel, k = 12, fill = NA, align = "right"),
      mean_wind_vel_1d = zoo::rollmean(wind_vel, k = 24 * 1, fill = NA, align = "right"),
      mean_wind_vel_3d = zoo::rollmean(wind_vel, k = 24 * 3, fill = NA, align = "right"),
      mean_wind_vel_5d = zoo::rollmean(wind_vel, k = 24 * 5, fill = NA, align = "right"),
      mean_rad_6h = zoo::rollmean(rad, k = 6, fill = NA, align = "right"),
      mean_rad_12h = zoo::rollmean(rad, k = 12, fill = NA, align = "right"),
      mean_rad_1d = zoo::rollmean(rad, k = 24 * 1, fill = NA, align = "right"),
      mean_rad_3d = zoo::rollmean(rad, k = 24 * 3, fill = NA, align = "right"),
      mean_rad_5d = zoo::rollmean(rad, k = 24 * 5, fill = NA, align = "right"),
      mean_humidity_6h = zoo::rollmean(humidity, k = 6, fill = NA, align = "right"),
      mean_humidity_12h = zoo::rollmean(humidity, k = 12, fill = NA, align = "right"),
      mean_humidity_1d = zoo::rollmean(humidity, k = 24 * 1, fill = NA, align = "right"),
      mean_humidity_3d = zoo::rollmean(humidity, k = 24 * 3, fill = NA, align = "right"),
      mean_humidity_5d = zoo::rollmean(humidity, k = 24 * 5, fill = NA, align = "right"),
      sum_precip_6h = zoo::rollsum(precip, k = 6, fill = NA, align = "right"),
      sum_precip_12h = zoo::rollsum(precip, k = 12, fill = NA, align = "right"),
      sum_precip_1d = zoo::rollsum(precip, k = 24 * 1, fill = NA, align = "right"),
      sum_precip_3d = zoo::rollsum(precip, k = 24 * 3, fill = NA, align = "right"),
      sum_precip_5d = zoo::rollsum(precip, k = 24 * 5, fill = NA, align = "right"),
      mean_water_temp_6h = zoo::rollmean(water_temp, k = 6, fill = NA, align = "right"),
      mean_water_temp_12h = zoo::rollmean(water_temp, k = 12, fill = NA, align = "right"),
      mean_water_temp_1d = zoo::rollmean(water_temp, k = 24 * 1, fill = NA, align = "right"),
      mean_water_temp_3d = zoo::rollmean(water_temp, k = 24 * 3, fill = NA, align = "right"),
      mean_water_temp_5d = zoo::rollmean(water_temp, k = 24 * 5, fill = NA, align = "right"),
      mean_water_temp_surface_6h = zoo::rollmean(water_temp_surface, k = 6, fill = NA, align = "right"),
      mean_water_temp_surface_12h = zoo::rollmean(water_temp_surface, k = 12, fill = NA, align = "right"),
      mean_water_temp_surface_1d = zoo::rollmean(water_temp_surface, k = 24 * 1, fill = NA, align = "right"),
      mean_water_temp_surface_3d = zoo::rollmean(water_temp_surface, k = 24 * 3, fill = NA, align = "right"),
      mean_water_temp_surface_5d = zoo::rollmean(water_temp_surface, k = 24 * 5, fill = NA, align = "right")) |>
    mutate_at(vars(starts_with("sum_precip")), ~ ifelse(. < 0.1, 0, .))
 
source("../R/roll_mean_wind_dir.R") 
 MeteoCH_data_hourly <- roll_mean_wind_dir(
  df = MeteoCH_data_hourly,
  dir_col = "wind_dir",
  speed_col = "wind_vel",
  k = c(6, 12, 24, 72, 120),
  suffix = c("6h", "12h", "1d", "3d", "5d")
)
```


# Join the hourly Thun network data with the other Meteo data
```{r}
all_data_hourly <- Thun_Network_data_hourly |>
  select(time, Log_NR, T_mean) |>
  rename(air_temp_LCD = T_mean) |>
  left_join(meta, by = "Log_NR") |>
  left_join(MeteoCH_data_hourly, by = "time") |>
  select(-X, -HuM, -ID) |>
  mutate(UHI = air_temp_LCD - air_temp)
```


# Aggregate the meteodata to daily nighttime data
```{r}
MeteoCH_data_daily_night <- MeteoCH_data_hourly |>
  filter(hour(time) >= 22 | hour(time) <= 6) |>
  # shift time 7h back
  mutate(time = time - hours(7)) |>
  group_by(date = as.Date(time)) |>
  summarise(air_temp = mean(air_temp, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE),
            precip_boolean = ifelse(precip > 0.3, 1, 0),
            RH_2m = mean(RH_2m_h, na.rm = TRUE),
            #glob_rad = mean(glob_rad_h, na.rm = TRUE),
            airpress = mean(airpress_h, na.rm = TRUE),
            CAP9 = first(CAP9),
            CAP18 = first(CAP18),
            CAP27 = first(CAP27),
            GWT8_Z500 = first(GWT8_Z500),
            GWT10_SLP = first(GWT10_SLP),
            GWT10_Z500 = first(GWT10_Z500),
            GWT18_SLP = first(GWT18_SLP),
            GWT18_Z500 = first(GWT18_Z500),
            GWT26_SLP = first(GWT26_SLP),
            GWT26_Z500 = first(GWT26_Z500)) |>
  ungroup()

# Daily global radiation
glob_rad_daily <- MeteoCH_data_hourly |>
  # shift time 7h back
  mutate(time = time - hours(7)) |>
  select(time, glob_rad_h) |>
  group_by(date = as.Date(time)) |>
  summarise(glob_rad = mean(glob_rad_h, na.rm = TRUE))

# Merge the two dataframes
MeteoCH_data_daily_night <- full_join(MeteoCH_data_daily_night, glob_rad_daily, by = "date")

wind_df <- MeteoCH_data_hourly |>
  filter(hour(time) >= 22 | hour(time) <= 6) |>
  # shift time 7h back
  mutate(time = time - hours(7)) |>
  select(time, wind_dir_h, wind_vel_h_kmh) |>
  rename(date = time,
         wd = wind_dir_h,
         ws = wind_vel_h_kmh)

# summarize wind speed and direction using function "openair" package
wind_df <- timeAverage(wind_df,
                         avg.time = "day",
                         statistic = "mean")

# Merge the two dataframes
MeteoCH_data_daily_night <- full_join(MeteoCH_data_daily_night, wind_df, by = "date") |>
    mutate(wd_class = case_when(
              wd < 22.5 & wd >= 337.5 ~ "N",
              wd >= 22.5 & wd < 67.5 ~ "NE",
              wd >= 67.5 & wd < 112.5 ~ "E",
              wd >= 112.5 & wd < 157.5 ~ "SE",
              wd >= 157.5 & wd < 202.5 ~ "S",
              wd >= 202.5 & wd < 247.5 ~ "SW",
              wd >= 247.5 & wd < 292.5 ~ "W",
              wd >= 292.5 & wd < 337.5 ~ "NW"),
           wd_NS = ifelse(wd > 90 & wd <= 270, "S", "N"),
           Wind_S = ifelse(wd_NS == "S", 1, 0)) |>
  rename(time = date)

# Add Thun Network data
all_data_daily_night <- full_join(MeteoCH_data_daily_night, Thun_Network_data_daily, by = "time")
write.csv(all_data_daily_night, "../data/final_predictors/meteo_data_daily_night.csv")
```

```{r}
# Combine the three dataframes: 10 minute data
all_data_10min <- full_join(MeteoCH_data_10min, Thun_Network_data, by = "time") |>
  left_join(meta, by = "Log_NR") |>
  left_join(Water_data, by = "time") 
write.csv(all_data_10min, "../data/final_predictors/all_data_10min.csv")

# hourly data
all_data_hourly <- full_join(MeteoCH_data_hourly, Thun_Network_data_hourly, by = "time") |>
  left_join(meta, by = "Log_NR") |>
  left_join(Water_data_hourly, by = "time") 
write.csv(all_data_hourly, "../data/final_predictors/all_data_hourly.csv")

# daily data night
all_data_daily_night <- full_join(MeteoCH_data_daily_night, Thun_Network_data_daily, by = "time") |>
  left_join(meta, by = "Log_NR") |>
  left_join(Water_data_daily_night, by = "time") 
write.csv(all_data_daily_night, "../data/final_predictors/all_data_daily_night.csv")
```

Create files with only the predictors
```{r}
predictors_hourly <- all_data_hourly |>
  select(-X, -T_mean, -T_min, -T_max, -HI_mean, -H_mean, -H_max, -H_min, -UHI_min, -UHI_max, -HI_min, -HI_max, -change_rate, -Location, -HuM, -LCZ_mixed, -LCZ, -LU_3, -LU_7, -ID)
write.csv(predictors_hourly, "../data/final_predictors/predictors_hourly.csv")

predictors_daily_night <- all_data_daily_night |>
  select(-X, -T_mean, -T_min, -T_max, -HI_mean, -H_mean, -H_max, -H_min, -UHI_min, -UHI_max, -HI_min, -HI_max, -Location, -HuM, -LCZ_mixed, -LCZ, -LU_3, -LU_7, -ID, -valid_hours, -UHI_mean, -valid_hours_night, -valid_hours_day, -T_min_night, -Tropical_night, -Gfroerli_night, -T_mean_night, -T_mean_day, -T_max_day, -Summer_day, -Hot_day, -Very_hot_day, -UHI_day)
write.csv(predictors_daily_night, "../data/final_predictors/predictors_daily_night.csv")
```


# Wind data analysis
First, we look at some wind roses.
```{r}
# Wind direction
MeteoCH_data_wind <- all_data_10min |>
  select(time, wind_dir_10min, wind_vel_10min_ms, wind_dir_10min_class, wind_dir_10min_NS, UHI, LU_7.type) |>
  drop_na()

ggwindrose(
  speed = MeteoCH_data_wind$wind_vel_10min_ms,
  direction = MeteoCH_data_wind$wind_dir_10min,
  speed_cuts = c(1, 2, 4, 8, 12),
  legend_title = "Wind speed (m/s)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
ggsave("../figures/Windroses/Wind_Rose_10min_data.png", width = 6, height = 5, dpi = 300)

# Nighttime wind
MeteoCH_data_wind_night <- MeteoCH_data_wind |>
  filter(hour(time) >= 22 | hour(time) <= 6)

ggwindrose(
  speed = MeteoCH_data_wind_night$wind_vel_10min_ms,
  direction = MeteoCH_data_wind_night$wind_dir_10min,
  speed_cuts = c(1, 2, 4, 8, 12),
  legend_title = "Wind speed (m/s)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose NIGHT (22pm-6am)",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
ggsave("../figures/Windroses/Wind_Rose_10min_data_night.png", width = 6, height = 5, dpi = 300)

# Daytime wind
MeteoCH_data_wind_day <- MeteoCH_data_wind |>
  filter(hour(time) >= 6 & hour(time) <= 22)
ggwindrose(
  speed = MeteoCH_data_wind_day$wind_vel_10min_ms,
  direction = MeteoCH_data_wind_day$wind_dir_10min,
  speed_cuts = c(1, 2, 4, 8, 12),
  #speed_cuts = c(1, 2, 3, 5, 12),
  legend_title = "Wind speed (m/s)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose DAY (6am-10pm)",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
ggsave("../figures/Windroses/Wind_Rose_10min_data_day.png", width = 6, height = 5, dpi = 300)
```

Plot windroses for different UHIs
```{r}
# filter for land use
MeteoCH_data_wind_UHI <- MeteoCH_data_wind |>
  filter(LU_7.type == "dense urban" | LU_7.type == "green urban" | LU_7.type == "rural") |>
  group_by(time) |>
  summarise(wind_dir_10min = mean(wind_dir_10min, na.rm = TRUE),
            wind_vel_10min_ms = mean(wind_vel_10min_ms, na.rm = TRUE),
            wind_dir_10min_NS = first(wind_dir_10min_NS),
            UHI = mean(UHI, na.rm = TRUE),
            LU_7.type = first(LU_7.type))

# Plot distribution of UHI
ggplot(MeteoCH_data_wind_UHI, aes(x = UHI)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of UHI",
       x = "UHI",
       y = "Frequency") +
  theme_minimal()

MeteoCH_data_wind_UHI_1 <- MeteoCH_data_wind_UHI |>
  filter(UHI > 1)

# plot the windrose
ggwindrose(
  speed = MeteoCH_data_wind_UHI_1$wind_vel_10min_ms,
  direction = MeteoCH_data_wind_UHI_1$wind_dir_10min,
  speed_cuts = c(1, 2, 4, 8, 12),
  legend_title = "Wind speed (m/s)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose UHI > 1",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)

MeteoCH_data_wind_UHI_neg <- MeteoCH_data_wind_UHI |>
  filter(UHI < 0)

# plot the windrose
ggwindrose(
  speed = MeteoCH_data_wind_UHI_neg$wind_vel_10min_ms,
  direction = MeteoCH_data_wind_UHI_neg$wind_dir_10min,
  speed_cuts = c(1, 2, 4, 8, 12),
  legend_title = "Wind speed (m/s)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose UHI < 0",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)

# plot windroses only for nighttime
MeteoCH_data_wind_UHI_1_night <- MeteoCH_data_wind_UHI_1 |>
  filter(hour(time) >= 22 | hour(time) <= 6)
ggwindrose(
  speed = MeteoCH_data_wind_UHI_1_night$wind_vel_10min_ms,
  direction = MeteoCH_data_wind_UHI_1_night$wind_dir_10min,
  speed_cuts = c(1, 2, 4, 8, 12),
  legend_title = "Wind speed (m/s)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose nightly UHI > 1 (22pm-6am)",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)

MeteoCH_data_wind_UHI_neg_night <- MeteoCH_data_wind_UHI_neg |>
  filter(hour(time) >= 22 | hour(time) <= 6)
ggwindrose(
  speed = MeteoCH_data_wind_UHI_neg_night$wind_vel_10min_ms,
  direction = MeteoCH_data_wind_UHI_neg_night$wind_dir_10min,
  speed_cuts = c(1, 2, 4, 8, 12),
  legend_title = "Wind speed (m/s)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose nightly UHI < 0 (22pm-6am)",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
```


# Correlate wind speed and direction with UHI
```{r}
# Correlate wind speed and direction with UHI
# symbol of the points should be according to wind_dir_10min_NS
ggplot(MeteoCH_data_wind_UHI, aes(x = wind_vel_10min_ms, y = UHI, color = wind_dir_10min_NS)) +
  geom_point(alpha = 0.4) +  # semi-transparent points
  geom_smooth(method = "lm", se = FALSE) +  # regression lines per group
  labs(
    title = "Correlation between Wind Speed and UHI",
    x = "Wind Speed (m/s)",
    y = "UHI"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("N" = "blue", "S" = "red"))
```

Do the whole thing again with hourly data
```{r}
# Wind direction
MeteoCH_data_wind <- all_data_hourly |>
  select(time, wind_dir_h, wind_vel_h_kmh, wind_dir_h_class, wind_dir_h_NS, UHI_mean, LU_7.type) |>
  drop_na()

ggwindrose(
  speed = MeteoCH_data_wind$wind_vel_h_kmh,
  direction = MeteoCH_data_wind$wind_dir_h,
  speed_cuts = c(2, 5, 10, 20),
  legend_title = "Wind speed (km/h)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)

# Nighttime wind
MeteoCH_data_wind_night <- MeteoCH_data_wind |>
  filter(hour(time) >= 22 | hour(time) <= 6)

ggwindrose(
  speed = MeteoCH_data_wind_night$wind_vel_h_kmh,
  direction = MeteoCH_data_wind_night$wind_dir_h,
  speed_cuts = c(2, 5, 10, 20),
  legend_title = "Wind speed (km/h)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose NIGHT (22pm-6am)",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)

# Daytime wind
MeteoCH_data_wind_day <- MeteoCH_data_wind |>
  filter(hour(time) >= 6 & hour(time) <= 22)
ggwindrose(
  speed = MeteoCH_data_wind_day$wind_vel_h_kmh,
  direction = MeteoCH_data_wind_day$wind_dir_h,
  speed_cuts = c(2, 5, 10, 20),
  legend_title = "Wind speed (km/h)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose DAY (6am-10pm)",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
```


```{r}
# filter for land use
MeteoCH_data_wind_UHI <- MeteoCH_data_wind |>
  filter(LU_7.type == "dense urban" | LU_7.type == "green urban" | LU_7.type == "rural") |>
  group_by(time) |>
  summarise(wind_dir_h = mean(wind_dir_h, na.rm = TRUE),
            wind_vel_h_kmh = mean(wind_vel_h_kmh, na.rm = TRUE),
            wind_dir_h_NS = first(wind_dir_h_NS),
            UHI = mean(UHI_mean, na.rm = TRUE),
            LU_7.type = first(LU_7.type))

MeteoCH_data_wind_UHI_1 <- MeteoCH_data_wind_UHI |>
  filter(UHI > 1)

# plot the windrose
ggwindrose(
  speed = MeteoCH_data_wind_UHI_1$wind_vel_h_kmh,
  direction = MeteoCH_data_wind_UHI_1$wind_dir_h,
  speed_cuts = c(2, 5, 10, 20),
  legend_title = "Wind speed (km/h)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose UHI > 1",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
ggsave("../figures/Windroses/Wind_Rose_UHIover1.png", width = 6, height = 5, dpi = 300)

MeteoCH_data_wind_UHI_neg <- MeteoCH_data_wind_UHI |>
  filter(UHI < 0)

# plot the windrose
ggwindrose(
  speed = MeteoCH_data_wind_UHI_neg$wind_vel_h_kmh,
  direction = MeteoCH_data_wind_UHI_neg$wind_dir_h,
  speed_cuts = c(2, 5, 10, 20),
  legend_title = "Wind speed (km/h)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose UHI < 0",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
ggsave("../figures/Windroses/Wind_Rose_UHIneg.png", width = 6, height = 5, dpi = 300)

MeteoCH_data_wind_UHI_low <- MeteoCH_data_wind_UHI |>
  filter(UHI > -0.3 & UHI < 0.3)

# plot the windrose
ggwindrose(
  speed = MeteoCH_data_wind_UHI_low$wind_vel_h_kmh,
  direction = MeteoCH_data_wind_UHI_low$wind_dir_h,
  speed_cuts = c(2, 5, 10, 20),
  legend_title = "Wind speed (km/h)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose low UHI (-0.3 < UHI < 0.3) ",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
ggsave("../figures/Windroses/Wind_Rose_UHIlow.png", width = 6, height = 5, dpi = 300)

# plot windroses only for nighttime
MeteoCH_data_wind_UHI_1_night <- MeteoCH_data_wind_UHI_1 |>
  filter(hour(time) >= 22 | hour(time) <= 6)
ggwindrose(
  speed = MeteoCH_data_wind_UHI_1_night$wind_vel_h_kmh,
  direction = MeteoCH_data_wind_UHI_1_night$wind_dir_h,
  speed_cuts = c(2, 5, 10, 20, 40),
  legend_title = "Wind speed (km/h)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose nightly UHI > 1 (22pm-6am)",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
ggsave("../figures/Windroses/Wind_Rose_UHIover1_night.png", width = 6, height = 5, dpi = 300)

MeteoCH_data_wind_UHI_neg_night <- MeteoCH_data_wind_UHI_neg |>
  filter(hour(time) >= 22 | hour(time) <= 6)
ggwindrose(
  speed = MeteoCH_data_wind_UHI_neg_night$wind_vel_h_kmh,
  direction = MeteoCH_data_wind_UHI_neg_night$wind_dir_h,
  speed_cuts = c(2, 5, 10, 20, 40),
  legend_title = "Wind speed (km/h)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose nightly UHI < 0 (22pm-6am)",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
ggsave("../figures/Windroses/Wind_Rose_UHIneg_night.png", width = 6, height = 5, dpi = 300)

# low UHI during night
MeteoCH_data_wind_UHI_low_night <- MeteoCH_data_wind_UHI_low |>
  filter(hour(time) >= 22 | hour(time) <= 6)

ggwindrose(
  speed = MeteoCH_data_wind_UHI_low_night$wind_vel_h_kmh,
  direction = MeteoCH_data_wind_UHI_low_night$wind_dir_h,
  speed_cuts = c(2, 5, 10, 20, 40),
  legend_title = "Wind speed (km/h)",
  calm_wind = 0,
  n_col = 1,
  plot_title = "Thun Wind Rose nightly low UHI (-0.3 < UHI < 0.3) (22pm-6am)",
) + labs(
  subtitle = "May-Oct 2024",
  caption = "Source: MeteoSwiss"
)
ggsave("../figures/Windroses/Wind_Rose_UHIlow_night.png", width = 6, height = 5, dpi = 300)
```

```{r}
# Correlate wind speed and direction with UHI
# symbol of the points should be according to wind_dir_10min_NS
ggplot(MeteoCH_data_wind_UHI, aes(x = wind_vel_h_kmh, y = UHI, color = wind_dir_h_NS)) +
  geom_point(alpha = 0.4) +  # semi-transparent points
  geom_smooth(method = "lm", se = FALSE) +  # regression lines per group
  labs(
    title = "Correlation between Wind Speed and UHI",
    x = "Wind Speed (km/h)",
    y = "UHI"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("N" = "purple", "S" = "darkcyan"))

# If we cut all above 10 km/h windspeed, the regression is different:
MeteoCH_data_wind_UHI |>
  filter(hour(time) >= 22 | hour(time) <= 6) |>
ggplot( aes(x = wind_vel_h_kmh, y = UHI, color = wind_dir_h_NS)) +
  geom_point(alpha = 0.4) +  # semi-transparent points
  geom_smooth(method = "lm", se = FALSE) +  # regression lines per group
  labs(
    title = "Correlation between Wind Speed and nighttime UHI",
    x = "Wind Speed (km/h)",
    y = "UHI"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("N" = "purple", "S" = "darkcyan"))

MeteoCH_data_wind_UHI |>
  filter(wind_vel_h_kmh < 10) |>
  filter(hour(time) >= 22 | hour(time) <= 6) |>
  ggplot(aes(x = wind_vel_h_kmh, y = UHI, color = wind_dir_h_NS)) +
  geom_point(alpha = 0.4) +  # semi-transparent points
  geom_smooth(method = "lm", se = FALSE) +  # regression lines per group
  labs(
    title = "Correlation between Wind Speed and nighttime UHI",
    x = "Wind Speed (km/h)",
    y = "UHI"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("N" = "purple", "S" = "darkcyan"))

MeteoCH_data_wind_UHI |> # seems like the hours are not really decisive
  filter(wind_vel_h_kmh < 10) |>
  filter(hour(time) >= 22 | hour(time) <= 6) |>
  ggplot(aes(x = wind_vel_h_kmh, y = UHI, color = as.factor(hour(time)))) +
  geom_point(alpha = 0.4) +  # semi-transparent points
  geom_smooth(method = "lm", se = FALSE) +  # regression lines per group
  labs(
    title = "Correlation between Wind Speed and nighttime UHI",
    x = "Wind Speed (km/h)",
    y = "UHI"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("22" = "#ffff55", "23" = "#ffcc00", "0" = "#ffaa11", "1" = "#ff8022", "2" = "#ff6622", "3" = "#ff5044", "4" = "#ee5077", "5" = "#dd4099", "6" = "magenta"))
```

Plot hourly UHI and wind direction
```{r}
p <- ggplot(MeteoCH_data_wind_UHI, aes(x = time, y = UHI)) +
  geom_line() +  # semi-transparent points
    # Custom the Y scales:
  scale_y_continuous(
    
    # Features of the first axis
    name = "UHI",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis( trans=~.*100, name="Wind direction")
  ) +
  # add wind direction on secondary y axis
  geom_point(aes(y = wind_dir_h/100), color = "#4466dd") +
  labs(
    title = "UHI and wind direction",
    x = "time",
  ) +
  theme_minimal()

plotly::ggplotly(p) |>
  layout(yaxis2 = list(overlaying = "y", side = "right", title = "Wind direction"))

```












