---
title: "Summary"
author: 'Moritz Burger & Noémie Wellinger'
date: "`r Sys.Date()`"
output: html_document
---

This Rmd script calculates important Urban Climate indicators for each logger. These are first calculated for each day of the measurement period, and finally summarized (averages) in a big summary table. Note that the time of day over they are calculated differs: whole day (00-24), daytime (6-22) and night (either 22-6 or 21-7).

\*\* NOT SURE HOW EXACTLY WE SHOULD PROCEED FOR THE VALID DAYS. THIS IS NOÉMIE'S SUGGESTION: **To contextualize the number of hot day for example, we also check the amount of valid days, as in no more than 20% NA. If there are more than 20% NA, all the data of this day is deleted.** 20% OF COURSE COULD ALSO BE 10%. ALSO, DOES IT MAKE SENSE FOR ALL INDICATORS TO USE THE SAME THRESHOLD? \*\*

```{r libraries, include=F}
library(tidyverse)
library(dplyr)
library(lubridate)
library(plotly)
library(weathermetrics)
```

# Preprocessing

```{r}
# Set variables
year <- 2024
# date_start = "2024-06-01"
# date_end = "2024-08-31"
date_start = "2024-05-01"
date_end = "2024-09-30"
date_start = as.Date(date_start, tz = "Europe/Zurich")
date_end = as.Date(date_end, tz = "Europe/Zurich")

city = "Thun"

variable = "T"
#variable = "H"

quality = "raw"
#quality = "QC"

ndays = as.numeric(difftime(as.Date(date_end)+1, as.Date(date_start), units = "days"))
ndays
```

```{r}
# Define Data and Metadata
data_T <- read_csv("../data/Thun Messnetz 2024/Logger_data_T_2024-05-01_2024-09-30.csv")
data_H <- read_csv("../data/Thun Messnetz 2024/Logger_data_H_2024-05-01_2024-09-30.csv")

meta <- read_csv('../data/Metadata/metadata_Thun.csv', show_col_types = FALSE)
meta_LU <- read_csv('../data/Metadata/Land_Use_Thun.csv', show_col_types = FALSE)

meta <- meta |>
  dplyr::select(Log_NR, STANDORT, NORD_CHTOPO, OST_CHTOPO, LV_03_N, LV_03_E) |>
  mutate(Log_NR = paste0("Log_", Log_NR)) |>
  left_join(meta_LU, by = c("STANDORT" = "Location")) |>
  rename(Log_NR = Log_NR.x) |>
  distinct() #remove duplicate rows
```

```{r}
# Check the time column
# if the first column of data is not called "time", change the name of the first column to "time"
if(colnames(data_T)[1] != "time"){
  colnames(data_T)[1] <- "time"
}

# Check the timezone of the time column
timezone <- attr(data_T$time, "tzone")
print(paste("Timezone is ", timezone))
```

Convert timezone if needed

```{r}
# get a list of Olson timezones
OlsonNames()
```

```{r}
# Set the current timezone, target timezone and time format
timezone <- "UTC" 
target_timezone <- "Europe/Zurich" 
format <- "%Y-%m-%d %H:%M"  # or %d.%m.%Y %H:%M or %Y-%m-%d %H:%M:%S or %d/%m/%Y %H:%M

# Convert:
data_T$time <- as.POSIXct(data_T$time, format = format, tz = timezone)
data_H$time <- as.POSIXct(data_H$time, format = format, tz = timezone)
# Shift to the target timezone
data_T$time <- with_tz(data_T$time, tzone = target_timezone)
data_H$time <- with_tz(data_H$time, tzone = target_timezone)
print(paste("Timezone is now set to", attr(data_T$time, "tzone"), "for data_T and", attr(data_H$time, "tzone"), "for data_H."))
```

```{r read, include = F}
# Shift the hours, so that the night is not cut in half
# cut the data to beginning and end date
data_T <- data_T |>
  filter(time >= date_start, time <= date_end) |>   # cut from date_start to date_end
  mutate(timeshift = time - dhours(7))  # 07 is the new 00
data_H <- data_H |>
  filter(time >= date_start, time <= date_end) |>
  mutate(timeshift = time - dhours(7))

visdat::vis_miss(
  data_T, #your table
  cluster = FALSE, 
  warn_large_data = FALSE
  )
visdat::vis_miss(
  data_H, #your table
  cluster = FALSE, 
  warn_large_data = FALSE
  )
```

# Summarize the data

## Long data format
```{r}
data_T_long <- data_T |>
  pivot_longer(cols = starts_with("Log_"), names_to = "Log_NR", values_to = "Temperature")

data_H_long <- data_H |>
  pivot_longer(cols = starts_with("Log_"), names_to = "Log_NR", values_to = "Humidity")

data_long <- data_T_long |>
  full_join(data_H_long, by = c("time", "timeshift", "Log_NR")) |>
  dplyr::select(time, timeshift, Log_NR, Temperature, Humidity)

# Calculate indices
data_long <- data_long |>
  # UHI
  left_join(# Create a reference table with Log_401's temperature at each time
    data_long |>
      filter(Log_NR == "Log_401")|>
      select(time, Temperature_401 = Temperature), 
    by = "time"
  ) |>
  mutate(UHI = Temperature - Temperature_401) |>
  select(-Temperature_401) |>
  # Heat Index
  mutate(Heat_Index = heat.index(t = Temperature, rh = Humidity, temperature.metric = "celsius", output.metric = "celsius", round = 2))
```

## Aggregate the data

### Hourly
```{r}
## Aggregate hourly, this may take long (like 2mins)
data_long_hourly <- data_long |>
  mutate(time_hour = floor_date(time, unit = "hour")) |> # round to nearest hour (floor)
  group_by(time_hour, Log_NR) |>
  summarise(
    T_mean = mean(Temperature, na.rm = TRUE),
    T_min = min(Temperature, na.rm = TRUE),
    T_max = max(Temperature, na.rm = TRUE),
    H_mean = mean(Humidity, na.rm = TRUE),
    H_min = min(Humidity, na.rm = TRUE),
    H_max = max(Humidity, na.rm = TRUE),
    UHI_mean = mean(UHI, na.rm = TRUE),
    UHI_min = min(UHI, na.rm = TRUE),
    UHI_max = max(UHI, na.rm = TRUE),
    HI_mean = mean(Heat_Index, na.rm = TRUE),
    HI_min = min(Heat_Index, na.rm = TRUE),
    HI_max = max(Heat_Index, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(across(everything(), ~ ifelse(is.nan(.) | is.infinite(.), NA, .))) |>
  mutate(time_hour = as.POSIXct(time_hour, origin = "1970-01-01", tz = target_timezone)) # Re-convert to POSIXct object

# recalculate mean UHI
data_long_hourly <- data_long_hourly |>
  # Re-alculate UHI from hourly means
  left_join(# Create a reference table with Log_401's temperature at each time
    data_long_hourly |>
      filter(Log_NR == "Log_401")|>
      select(time_hour, Temperature_401 = T_mean), 
    by = c("time_hour")
  ) |>
  mutate(UHI_mean = T_mean - Temperature_401) |>
  select(-Temperature_401)
  # Re-Calculate Heat Index
  #mutate(Heat_Index_hourly = heat.index(t = Temperature, rh = Humidity, temperature.metric = "celsius", output.metric = "celsius", round = 2)) 
  # Checkt if there is a difference between the hourly and the daily calculation - there is one
  #mutate(UHI_diff = round(UHI - UHI_hourly, 2), Heat_Index_diff = round(Heat_Index - Heat_Index_hourly, 2))
```

### Diurnal
```{r}
data_long_diurnal_MJJAS <- data_long_hourly |>
  mutate(hour=hour(time_hour)) |>
  filter(month(time_hour) %in% c(5:9)) |>
  group_by(hour, Log_NR) |>
  summarise(across(c(T_mean, H_mean, UHI_mean, HI_mean), ~ mean(., na.rm = TRUE)),
            .groups = "drop") |>
  mutate(across(everything(), ~ ifelse(is.nan(.)|is.infinite(.), NA, .)))

data_long_diurnal_JAS <- data_long_hourly |>
  mutate(hour=hour(time_hour)) |>
  filter(month(time_hour) %in% c(7:9)) |>
  group_by(hour, Log_NR) |>
  summarise(across(c(T_mean, H_mean, UHI_mean, HI_mean), ~ mean(., na.rm = TRUE)),
            .groups = "drop") |>
  mutate(across(everything(), ~ ifelse(is.nan(.)|is.infinite(.), NA, .)))
```

### Daily
```{r}
## Aggregate daily
### 1. Data aggregated from hourly data: Mean values. This compensates for missing 10-min data points if some hours have more NA values than others.
data_long_daily2 <- data_long_hourly |>
  mutate(time_day = floor_date(time_hour, unit = "day")) |> # round to day (floor)
  group_by(time_day, Log_NR) |>
  summarise(valid_hours = sum(!is.na(T_mean)), # Calculate valid hours as number of hours per day that contain data (not NA)
            across(c(T_mean, H_mean, UHI_mean, HI_mean), ~ mean(., na.rm = TRUE)),
            across(c(T_min, H_min, UHI_min, HI_min), ~ min(., na.rm = TRUE)),
            across(c(T_max, H_max, UHI_max, HI_max), ~ max(., na.rm = TRUE)),
            .groups = "drop") |>
  mutate(across(everything(), ~ ifelse(is.nan(.)|is.infinite(.), NA, .))) |>
  mutate(time_day = as.POSIXct(time_day, origin = "1970-01-01", tz = target_timezone)) # Re-convert to POSIXct object


### 2. Re-calculate UHI from daily means
data_long_daily3 <- data_long_daily2 |>
  # re-calculate UHI
  left_join(# Create a reference table with Log_401's temperature at each time
    data_long_daily2 |>
      filter(Log_NR == "Log_401")|>
      select(time_day, Temperature_401 = T_mean), 
    by = c("time_day")
  ) |>
  mutate(UHI_mean = T_mean - Temperature_401) |>
  select(-Temperature_401) 


### 3. Nighttime values and indicators (21:00 - 07:00): Tmin, Tropical Nights, Gfrörli Nights
data_long_daily_night1 <- data_long_hourly |>
  mutate(timeshift = time_hour - dhours(7),
         time_day = floor_date(time_hour, unit = "day")) |> # 07 is the new 00
  filter(hour(timeshift) >= 14) |>
  group_by(time_day, Log_NR) |>
  summarise(T_min_night = min(T_min, na.rm = TRUE),
            .groups = "drop") |>
  mutate(across(everything(), ~ ifelse(is.nan(.)|is.infinite(.), NA, .)),
         time_day = as.POSIXct(time_day, origin = "1970-01-01", tz = target_timezone)) |> # Re-convert to POSIXct object
  # Add indicators
  mutate(Tropical_night = ifelse(T_min_night >= 20 & T_min_night < 30, 1, 0),
         Gfroerli_night = ifelse(T_min_night <= 10 & T_min_night >= 0, 1, 0)) 


### 4. Nighttime values and indicators (22:00 - 06:00): T_mean, nighttime UHI
data_long_daily_night2 <- data_long_hourly |>
  mutate(timeshift = time_hour - dhours(6),
         time_day = floor_date(time_hour, unit = "day")) |> # 07 is the new 00
  filter(hour(timeshift) >= 16) |>
  group_by(time_day, Log_NR) |>
  summarise(T_mean_night = mean(T_mean, na.rm = TRUE),
            .groups = "drop") |>
  mutate(across(everything(), ~ ifelse(is.nan(.)|is.infinite(.), NA, .)),
         time_day = as.POSIXct(time_day, origin = "1970-01-01", tz = target_timezone)) # Re-convert to POSIXct object

# Re-calculate UHI from T_mean_night
data_long_daily_night2 <- data_long_daily_night2 |>
  left_join(# Create a reference table with Log_401's temperature at each time
    data_long_daily_night2 |>
      filter(Log_NR == "Log_401")|>
      select(time_day, Temperature_401 = T_mean_night), 
    by = c("time_day")
  ) |>
  mutate(UHI_night = T_mean_night - Temperature_401) |>
  select(-Temperature_401)


### 5. Daytime values and indicators (06:00 - 22:00): T_mean, daytime UHI
data_long_daily_day <- data_long_hourly |>
  filter(hour(time_hour) >= 6, hour(time_hour) < 22) |>
  mutate(time_day = floor_date(time_hour, unit = "day")) |> # 07 is the new 00
  group_by(time_day, Log_NR) |>
  summarise(T_mean_day = mean(T_mean, na.rm = TRUE),
            T_max_day = max(T_max, na.rm = TRUE),
            .groups = "drop") |>
  mutate(across(everything(), ~ ifelse(is.nan(.)|is.infinite(.), NA, .)),
         time_day = as.POSIXct(time_day, origin = "1970-01-01", tz = target_timezone)) |> # Re-convert to POSIXct object
  # Add daytime indicators
  mutate(Summer_day = ifelse(T_max_day >= 25 & T_max_day < 30, 1, 0),
         Hot_day = ifelse(T_max_day >= 30 & T_max_day < 35, 1, 0),
         Very_hot_day = ifelse(T_max_day >= 35, 1, 0))

# Re-calculate UHI from T_mean_day
data_long_daily_day <- data_long_daily_day |>
  left_join(# Create a reference table with Log_401's temperature at each time
    data_long_daily_day |>
      filter(Log_NR == "Log_401")|>
      select(time_day, Temperature_401 = T_mean_day), 
    by = c("time_day")
  ) |>
  mutate(UHI_day = T_mean_day - Temperature_401) |>
  select(-Temperature_401)
```

### Monthly

```{r}
# Aggregate monthly
data_long_monthly <- data_long_daily |>
  group_by(month, Log_NR) |>
  summarise(across(c(T_mean, H_mean, UHI_mean, HI_mean), ~ mean(., na.rm = TRUE)),
            across(c(T_min, H_min, UHI_min, HI_min), ~ min(., na.rm = TRUE)),
            across(c(T_max, H_max, UHI_max, HI_max), ~ max(., na.rm = TRUE)),
            across(c(Tropical_night, Gfroerli_night, Summer_day, Hot_day, Very_hot_day), ~ sum(., na.rm = TRUE)),
            .groups = "drop") |>
  mutate(across(everything(), ~ ifelse(is.nan(.)|is.infinite(.), NA, .)))
```

### Season

```{r}
# Aggregate over whole time period
data_long_MJJAS <- data_long_daily |>
  filter(month %in% c(5, 6, 7, 8, 9)) |>
  group_by(Log_NR) |>
  summarise(across(c(T_mean, H_mean, UHI_mean, HI_mean), ~ mean(., na.rm = TRUE)),
            across(c(T_min, H_min, UHI_min, HI_min), ~ min(., na.rm = TRUE)),
            across(c(T_max, H_max, UHI_max, HI_max), ~ max(., na.rm = TRUE)),
            across(c(Tropical_night, Gfroerli_night, Summer_day, Hot_day, Very_hot_day), ~ sum(., na.rm = TRUE)),
            .groups = "drop") |>
  mutate(across(everything(), ~ ifelse(is.nan(.)|is.infinite(.), NA, .)))

data_long_JAS <- data_long_daily |>
  filter(month %in% c(7, 8, 9)) |>
  group_by(Log_NR) |>
  summarise(across(c(T_mean, H_mean, UHI_mean, HI_mean), ~ mean(., na.rm = TRUE)),
            across(c(T_min, H_min, UHI_min, HI_min), ~ min(., na.rm = TRUE)),
            across(c(T_max, H_max, UHI_max, HI_max), ~ max(., na.rm = TRUE)),
            across(c(Tropical_night, Gfroerli_night, Summer_day, Hot_day, Very_hot_day), ~ sum(., na.rm = TRUE)),
            .groups = "drop") |>
  mutate(across(everything(), ~ ifelse(is.nan(.)|is.infinite(.), NA, .)))
```

## Mean values

```{r}
# Aggregate the data to hourly values
T_mean_hourly <- data_T |>
  mutate(month = month(time), 
         day = day(time), 
         hour = hour(time)) |>
  group_by(month, day, hour) |>
  summarise(across(everything(),~ mean(., na.rm = TRUE))) |>
  ungroup()

H_mean_hourly <- data_H |>
  mutate(month = month(time), 
         day = day(time), 
         hour = hour(time)) |>
  group_by(month, day, hour) |>
  summarise(across(everything(), ~ mean(., na.rm = TRUE))) |>
  ungroup()

T_mean_diurnal <- T_mean_hourly |>
  group_by(hour) |>
  summarise(across(starts_with("Log_"), ~ mean(., na.rm = TRUE)) ) |>
  ungroup()

H_mean_diurnal <- H_mean_hourly |>
  group_by(hour) |>
  summarise(across(starts_with("Log_"), ~ mean(., na.rm = TRUE)) ) |>
  ungroup()

T_mean_daily <- T_mean_hourly |>
  group_by(month, day) |>
  summarise(across(everything(), ~ mean(., na.rm = TRUE))) |>
  ungroup()

H_mean_daily <- H_mean_hourly |>
  group_by(month, day) |>
  summarise(across(everything(), ~ mean(., na.rm = TRUE))) |>
  ungroup()
```

## Indices

UHI, Heat index

```{r}
# UHI
UHI_10min <- data_T |>
  mutate(across(starts_with("Log_"), ~ .-Log_401))

UHI_hourly <- UHI_10min |>
  group_by(month = month(time), day = day(time), hour = hour(time)) |>
  summarise(across(starts_with("Log_"), function(x) if (all(is.na(x))) NA else mean(x, na.rm = TRUE)), # use ifelse to avoid NaN if all NA
            time = as.POSIXct(time, format = "%Y-%m-%d %H"),
            timeshift = as.POSIXct(timeshift, format = "%Y-%m-%d %H")) |>
  ungroup()

UHI_diurnal <- UHI_hourly |>
  group_by(hour) |>
  summarise(across(starts_with("Log_"), ~ mean(., na.rm = TRUE)) ) |>
  ungroup() |>
  dplyr::select(-Log_401)

UHI_daily <- UHI_hourly |>
  group_by(month = month(time), day = day(time)) |>
  summarise(across(starts_with("Log_"), function(x) if (all(is.na(x))) NA else mean(x, na.rm = TRUE)), # use ifelse to avoid NaN if all NA
            time = as.POSIXct(time, format = "%Y-%m-%d %H")) |>
  ungroup()

# Heat index



```

## Indicators using Nighttime Temperatures (21:00 - 07:00)

Tmin, Tropical Nights, Gfrörli Nights

```{r indicators_0, include = F}
data_T_night <- data_T |> 
  filter(hour(timeshift) >= 14)

Tmin_night <- data_T_night |> #daily
  group_by(month=month(timeshift), day=day(timeshift)) |>
  summarise_all(~ min(., na.rm = TRUE))
Tmin_night[sapply(Tmin_night, is.infinite)] <- NA

Tmin_night_summer <- Tmin_night |> #average over whole time period
  group_by(year = year(timeshift)) |>
  summarise_all(~ mean(., na.rm = TRUE)) |>
  dplyr::select(-time, -timeshift, -month, -day, -year)

Tropical_nights <- Tmin_night |> #daily
  group_by(month=month(timeshift), day=day(timeshift)) |>
  summarise_all(~ ifelse(. >= 20 & . < 30, 1, 0))

Tropical_nights_total <- Tropical_nights |> #average over whole time period
  group_by(time) |>
  summarise_all(~ sum(.), na.rm = TRUE) |>
  dplyr::select(-time, -timeshift, -month, -day)

Gfroerli_nights <- Tmin_night |> #daily
  group_by(month=month(timeshift), day=day(timeshift)) |>
  summarise_all(~ ifelse(. <=10 & . >= 0, 1, 0))

Gfroerli_nights_total <- Gfroerli_nights |> #average over whole time period
  group_by(time) |>
  summarise_all(~ sum(.), na.rm = TRUE) |>
  dplyr::select(-time, -timeshift, -month, -day)
```

## Indicators using Nighttime Temperatures (22:00 - 06:00)

Mean night temperature, nighttime UHI

```{r indicators_1, include=F}
data_T_night2 <- data_T |> 
  filter(hour(timeshift) >= 15, hour(timeshift)<23)

# mean night temperature
Tmean_night <- data_T_night2 |> #daily
  group_by(month=month(timeshift), day=day(timeshift)) |>
  summarise_all(~ mean(., na.rm = TRUE))
Tmean_night[sapply(Tmean_night, is.infinite)] <- NA
Tmean_night[sapply(Tmean_night, is.nan)] <- NA

Tmean_night_summer <- Tmean_night |>
  group_by(year = year(timeshift)) |>
  summarise_all(~ mean(., na.rm = TRUE)) |>
  dplyr::select(-time, -timeshift, -month, -day, -year)

# Urban heat island
# Reference: Meteoschweiz 3m (Log_401)
UHI_night <- Tmean_night |>
  mutate(across(everything(), ~ . - Log_401))
UHI_night[sapply(UHI_night, is.nan)] <- NA

UHI_night_avg <- UHI_night |>
  group_by(year = year(timeshift)) |>
  summarise_all(~ mean(., na.rm = TRUE)) |>
  dplyr::select(-time, -timeshift, -month, -year, -day) |>
  head(1) #select the first row, since there is a second row with NaN
```

# Indicators using daytime Temperatures (06:00 - 22:00)

Mean daytime temperature

```{r indicators_2, include=F}
data_day <- data |> 
  filter(hour(time) >= 6, hour(time) <= 22) # use time again, to avoid breaking the day apart

# mean day temperature
Tmean_day <- data_day |>
  group_by(month=month(time), day=day(time)) |>
  summarise_all(~ mean(., na.rm = TRUE))

Tmean_day[sapply(Tmean_day, is.infinite)] <- NA
Tmean_day[sapply(Tmean_day, is.nan)] <- NA

Tmean_day_avg <- Tmean_day |>
  group_by(year = year(time)) |>
  summarise_all(~ mean(., na.rm = TRUE)) |>
  dplyr::select(-time, -timeshift, -month, -day, -year)

# Urban heat island
# Reference: Zollikofen Referenz 3m (Log_99)
UHI_day <- Tmean_day |>
  mutate_all(~ .-Log_99) |>
  dplyr::select(-month, -day)
UHI_day[sapply(UHI_day, is.nan)] <- NA

UHI_day_avg <- UHI_day |>
  group_by(year = year(time)) |>
  summarise_all(~ mean(., na.rm = TRUE)) |>
  dplyr::select(-time, -timeshift, -month, -year) |>
  head(1) #select the first row, since there is a second row with NaN
```

# Indicators using measurements of entire day

```{r indicators_3, include = F}
# Overall mean temperature
Tmean <- data |>
  group_by(month=month(time), day=day(time)) |>
  summarise_all(~ mean(., na.rm = TRUE))

Tmean[sapply(Tmean, is.infinite)] <- NA
Tmean[sapply(Tmean, is.nan)] <- NA

Tmean_avg <- Tmean |>
  group_by(year = year(timeshift)) |>
  summarise_all(~ mean(., na.rm = TRUE)) |>
  dplyr::select(-time, -timeshift, -month, -day, -year)

# Overall max temperature
Tmax <- data |>
  group_by(month=month(time), day=day(time)) |>
  summarise_all(~ max(., na.rm = TRUE))

Tmax[sapply(Tmax, is.infinite)] <- NA
Tmax[sapply(Tmax, is.nan)] <- NA

Tmax_avg <- Tmax |>
  group_by(year = year(time)) |>
  summarise_all(~ mean(., na.rm = TRUE)) |>
  dplyr::select(-time, -timeshift, -month, -day, -year)

# Urban heat island
# Reference: Zollikofen Referenz 3m (Log_99)
UHI <- Tmean |>
  mutate_all(~ .-Log_99) |>
  dplyr::select(-month, -day)
UHI[sapply(UHI, is.nan)] <- NA

UHI_avg <- UHI |>
  group_by(year = year(time)) |>
  summarise_all(~ mean(., na.rm = TRUE)) |>
  dplyr::select(-time, -timeshift, -month, -year) |>
  head(1) #select the first row, since there is a second row with NaN

# Summer days
Summer_days <- Tmax |>
  dplyr::select(-timeshift, -month, -day) |>
  group_by(year = year(time)) |>
  summarise_all(~ sum(. >= 25 & . <= 50, na.rm = TRUE)) |>
  dplyr::select(-year, -month)

# Hot days
Hot_days <- Tmax |>
  dplyr::select(-timeshift, -month, -day) |>
  group_by(year = year(time)) |>
  summarise_all(~ sum(. >= 30 & . <= 50, na.rm = TRUE)) |>
  dplyr::select(-year, -month)

# Very Hot days
VHot_days <- Tmax |>
  dplyr::select(-timeshift, -month, -day) |>
  group_by(year = year(time)) |>
  summarise_all(~ sum(. >= 35 & . <= 50, na.rm = TRUE)) |>
  dplyr::select(-year, -month)
```

# Valid days

```{r valid_days, include = F}
# TODO
# valid days
# question also valid nights, and daytime ?
# when is a day valid? does it have to be completely without NA or not more than 20% NA or so?
# or also look at each hour
# gap filling with new sensors
# valid hours = 4 measurements or more per hour

# Set threshold for maximum amount of NA data that is allowed per day (e.g. 0.2 = 20%)
threshold <- 0.2

# Calculate the number of NA values per day
NA_per_day <- data |>
  group_by(date = date(time)) |>
  summarize_all(~ sum(is.na(.))) |>
  dplyr::select(-time)

# Flag the valid (1) and invalid (0) days
valid_days <- NA_per_day |>
  mutate(across(-date, ~ifelse(. > 144*threshold, 0, 1))) # 144 measurements per day (every 10 mins)

# Calculate the total number of days and valid days
total_valid_days <- valid_days |>
  dplyr::select(-date) |>
  summarise_all(~ sum(., na.rm = TRUE)) 
```

# Final summary table

```{r transpose, include = F}
# Transpose all the indicator tables
Tmean_avg <- as.data.frame(t(head(Tmean_avg))) |> rename(Tmean = V1) |> rownames_to_column(var = "Log_name")
Tmax_avg <- as.data.frame(t(head(Tmax_avg))) |> rename(Tmax = V1) |> rownames_to_column(var = "Log_name")
Tmean_night_avg <- as.data.frame(t(head(Tmean_night_avg))) |> rename(Tmean_night = V1) |> rownames_to_column(var = "Log_name")
Tmin_night_avg <- as.data.frame(t(head(Tmin_night_avg))) |> rename(Tmin_night = V1) |> rownames_to_column(var = "Log_name")
Tmean_day_avg <- as.data.frame(t(head(Tmean_day_avg))) |> rename(Tmean_day = V1) |> rownames_to_column(var = "Log_name")
UHI_avg <- as.data.frame(t(head(UHI_avg))) |> rename(UHI = V1) |> rownames_to_column(var = "Log_name")
UHI_day_avg <- as.data.frame(t(head(UHI_day_avg))) |> rename(UHI_day = V1) |> rownames_to_column(var = "Log_name")
UHI_night_avg <- as.data.frame(t(head(UHI_night_avg))) |> rename(UHI_night = V1) |> rownames_to_column(var = "Log_name")
Tropical_nights <- as.data.frame(t(head(Tropical_nights))) |> rename(Tropical_nights = V1) |> rownames_to_column(var = "Log_name")
Gfroerli_nights <- as.data.frame(t(head(Gfroerli_nights))) |> rename(Gfroerli_nights = V1) |> rownames_to_column(var = "Log_name")
Summer_days <- as.data.frame(t(head(Summer_days))) |> rename(Summer_days = V1) |> rownames_to_column(var = "Log_name")
Hot_days <- as.data.frame(t(head(Hot_days))) |> rename(Hot_days = V1) |> rownames_to_column(var = "Log_name")
VHot_days <- as.data.frame(t(head(VHot_days))) |> rename(VHot_days = V1) |> rownames_to_column(var = "Log_name")

valid_days <- as.data.frame(t(head(total_valid_days))) |> rename(total_valid_days = V1) |> rownames_to_column(var = "Log_name")

# Keep in mind this overwrites the original tables, so if you run this chunk again, it transposes back and probably causes an error

# Calculate also the percentage of valid days
valid_days <- valid_days |>
  mutate(valid_days_percent = total_valid_days/ndays)
```

```{r join, include=F}
# Join the transposed tables
indicators <- list(Tmean_avg, Tmax_avg, Tmean_night_avg, Tmin_night_avg, Tmean_day_avg, UHI_avg, UHI_day_avg, UHI_night_avg, Tropical_nights, Gfroerli_nights, Summer_days, Hot_days, VHot_days, valid_days)

Summary <- plyr::join_all(indicators, by = "Log_name", type = "full")
```

Add the coordinates from the metadata and format them correctly. Important for mapping later on.

```{r coordinates, include = F}
Summary <- Summary |>
  full_join(meta, by = 'Log_name') |>
  drop_na() |>
  distinct() #Remove identical rows
```

```{r}
# Remove stations that don't belong in the mean
Summary_all <- Summary |>
  # Remove Zollikofen 2m because we only want to map our 3m stations and stations with large errors or NAs
    dplyr::filter(Location != ("Zollikofen 2m")) |>
    dplyr::filter(Location != ("Ostermundigen BareTower")) |>
    filter(!grepl("Postgasse", Location, ignore.case = TRUE)) # remove all Postgasse Loggers

# Calculate mean over all stations
Summary_all <- Summary_all |>
  summarise_all(~ mean(., na.rm = TRUE)) 
```

```{r export, includ = F}
write_csv(Summary, (paste0("../data/", city, "/Summary_T_", quality, "_", date_start, "_", date_end, ".csv")))
```
